---
title: "Shrub invasion in forests - Analysis and figures"
author: "Laís Petri"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  html_document:
    toc: true
    theme: united 
    highlight: tango 
    toc_float: true 
editor_options: 
  chunk_output_type: console
---

**Please, read these notes before running the code:**

1. You will need to download and install OpenBUGS 
(https://www.mrc-bsu.cam.ac.uk/software/bugs/openbugs/)
2. The order presented here is defined by each question asked in the main paper. 
3. Analysis are performed first, then the respective figure.
4. Analysis and figures presented in the Supplementary material (SM) are placed at the end of this code, after
the main questions are tackled.
5. You will need to create the following folders in the working directory: "data", "figures", and "parameters".
6. Paste the files from this webpage (https://doi.org/10.5061/dryad.msbcc2g33) inside of the "data" folder you just created.
7. Note: figures included in the main manuscript or SM will not be displayed in the html file after knitting. They will be inside of the folder you created, "figures". The aspect ratio could not be maintained when knitting (probably due to my own coding limitation).

```{r setup, include=FALSE}
# suppress warnings and messages in case you decide to knit this file
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r packages, include=FALSE}
# loading the packages
## linked to many functions
require(tidyverse)
require(readxl)

## basic ones necessary for map
require(sf)

## map of the countries of the entire world and choosing scale, respectively
require(rnaturalearth)
require(rnaturalearthdata)
require(rgeos)
require(ggspatial)
require(ggthemes)

## managing data & graphing
require(ggrepel)
require(gridExtra)
require(ggpubr)
require(ggh4x)

## run models
require(R2OpenBUGS)
require(mcmcplots)

## graphs
require(pals) 
require(patchwork)
require(egg) 

# setting common theme
theme_set(theme_bw()) # classic dark-on-light theme
```

```{r directory, include=FALSE}
# Setting up paths to retrieve data files and outputs
workingdir <- getwd()
graph.wd <- file.path(workingdir, "figures")
datain <- file.path(workingdir, "data")
parametersout <- file.path(workingdir, "parameters")
```

```{r include=FALSE}
# Defining colors that will be used later on in the figures
## mechanisms
x<-as.data.frame(pals::brewer.dark2(3))
BRcolor <- x[3,]
ENcolor <- x[2,]
PPcolor <- x[1,]

## abiotic factors
x<-as.data.frame(pals::brewer.paired(12))
Lightcolor <- x[11,]
Nutrientcolor <- x[12,]
Watercolor <- x[10,]

## number of invaders
Singlecolor <- x[2,]
Multicolor <- x[8,]

## type of ES
ESNatcolor <- x[5,]
ESInvcolor <- x[9,]
```

# **Importing data**
```{r, echo=FALSE}
# Importing the data 
data <- read.csv(file.path(datain, "Data.csv"))
# Effect size (ES) estimates
InvEstimate <- read.csv(file.path(datain, "InvEstimateRESULTS.csv"))
NatBioticEstimate <- read.csv(file.path(datain, "NatBioticEstimateRESULTS.csv"))
NatAbioticEstimate <- read.csv(file.path(datain, "NatAbioticEstimateRESULTS.csv"))
```

Column names are reported in the supplementary material
```{r importing the estimated data, include=TRUE}
glimpse(data) # all data
glimpse(InvEstimate) # Effect size (ES) estimates for invasive species performance with variance
glimpse(NatBioticEstimate) # Effect size (ES) estimates for native community performance with variance
glimpse(NatAbioticEstimate) # Effect size (ES) estimates for abiotic conditions with variance
```

```{r, include=FALSE}
# Selecting unique observations
## Invasive ES
InvUnique <- data %>% select(ObsIDOriginal, InvContmean, InvContsd, InvTreatmean, InvTreatsd) %>% distinct(InvContmean, InvContsd, InvTreatmean, InvTreatsd, .keep_all = T) %>% select(ObsIDOriginal)

## Native community ES
NatUniqueBiotic <- data %>% filter(NativeCommunity == 1) %>% select(ObsIDOriginal, InvContmean, InvContsd, InvTreatmean, InvTreatsd, NatContmean, NatContsd, NatTreatmean, NatTreatsd) %>% distinct(InvContmean, InvContsd, InvTreatmean, InvTreatsd, NatContmean, NatContsd, NatTreatmean, NatTreatsd, .keep_all = T) %>% select(ObsIDOriginal) 

## Abiotic conditions ES
NatUniqueAbiotic <- data %>% filter(NativeResponse_CATEGORY == "abiotic", ObsIDOriginal != 1002, ObsIDOriginal != 1121, ObsIDOriginal != 1122) %>% select(ObsIDOriginal, InvContmean, InvContsd, InvTreatmean, InvTreatsd, NatContmean, NatContsd, NatTreatmean, NatTreatsd) %>% distinct(InvContmean, InvContsd, InvTreatmean, InvTreatsd, NatContmean, NatContsd, NatTreatmean, NatTreatsd, .keep_all = T) %>% select(ObsIDOriginal)
```

# **Q1: What is the most successful mechanism of invasion? What is the most impactful mechanism of invasion?**

## Invasive species performance
### data preparation
```{r echo=TRUE}
# here, I merge the file with the estimated ES values with SD and calculate ES for the observations with no associated variability
InvAllDataModel <- data %>% filter(ObsIDOriginal %in% InvUnique$ObsIDOriginal) %>% left_join(InvEstimate, by = "ObsIDOriginal") %>% mutate(ESInvmean = ifelse(is.na(ESInvSD), ((InvTreatmean-InvContmean)/((InvTreatmean+InvContmean)/2)), ESInvmean), ObsID = row_number()) %>% group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()
```

### analysis
#### organizing model data
```{r echo=TRUE}
J <- nrow(InvAllDataModel) # number of observations
StudyID <- InvAllDataModel$StudyID # data on study unique number (ID)
Mechanism <- InvAllDataModel$Mechanism # data on mechanisms
invESavemean <- InvAllDataModel$ESInvmean # data on ES mean
invESavesd <- InvAllDataModel$ESInvSD # data on ES variability
M <- max(InvAllDataModel$ESInvSD, na.rm = T) # maximum value of variability
K <- length(unique(InvAllDataModel$Mechanism)) # number of unique mechanisms
R <- K+1 # total number of tau
W <- length(unique(InvAllDataModel$StudyID)) # number of unique study IDs

# data file to run model
datamodel <- list (J=J, StudyID=StudyID, Mechanism=Mechanism, 
                   invESavemean=invESavemean, invESavesd=invESavesd, M=M, K=K, R=R, W=W)
```

#### model 
```{r echo=TRUE}
modelESInv <- function() {
  for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(invESavesd[i]*invESavesd[i])

invESavemean[i]~dnorm(iE[i],tau[Mechanism[i]])
invESavemean.h[i]~dnorm(iE[i],tau[Mechanism[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[Mechanism[i]]+iSRE[StudyID[i]]+ie[i]
residuals[i]<-invESavemean.h[i]-invESavemean[i]
}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

  for(i in 1:W){
iSRE[i]~dnorm(0,tau[R])
}

  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}
} # end model

# initials
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "invESavemean.h", "tau", "var", "residuals")

# running model
es.simInvDiffStudy <- bugs(datamodel, inits, parameters, model.file = modelESInv, 
                           n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simInvDiffStudy[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simInvDiffStudy, "beta")
```

+ Model parameters
```{r parameters 1, echo=FALSE}
x<-as.data.frame(es.simInvDiffStudy[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESInv-Mech.csv"), row.names = F)
x
```

+ Model residuals
```{r residuals 1, include=FALSE}
resInv<-as.data.frame(es.simInvDiffStudy[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^residuals", rowname))

write.csv(resInv, file.path(parametersout, "ESInv-MechRes.csv"), row.names = F)
```

+ Model fit:
```{r echo=FALSE}
x<-as.data.frame(es.simInvDiffStudy[["summary"]]) # retrieves model results
invESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^invESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(invESavemean.h,invESavemean)
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESInvmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 1.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation for plots of invasive ES
samplesize <- InvAllDataModel %>% 
  group_by(Mechanism) %>%
  add_tally() %>% ungroup() %>% select(Mechanism, n) %>% distinct() %>%
  arrange(Mechanism) %>%
  mutate(comparison2 = paste0("mechanism", 1:3))

         
inv <- as.data.frame(es.simInvDiffStudy[["summary"]])
inv <- inv %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(!grepl("^invESavemean.h", rowname),
         !grepl("^ESIA", rowname),
         !grepl("^tau", rowname),
         !grepl("^var",rowname),
         !grepl("^residuals",rowname),
         rowname != "deviance") %>%
  mutate(type = "invasive")

forestinv <- inv %>%
  arrange(rowname) %>%
  mutate(test = str_sub(rowname, 6, 6)) %>% 
  # mutate_all(na_if,"") %>% 
  mutate(comparison2 = ifelse(is.na(test), "overall", 
                             ifelse(!is.na(test), paste0('mechanism', test), NA))) %>%
  arrange(comparison2) %>%
  mutate(index = row_number()) %>%
  left_join(samplesize) %>%
  group_by(type) %>%
  mutate(n1 = ifelse(is.na(n), sum(n, na.rm=TRUE), n),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')'),
         index=ifelse(Mechanism==1, 3, ifelse(Mechanism==2, 1, 2))) %>%
  select(-n1, -test) %>%
  ungroup()
```

```{r include=FALSE}
# adjusts the x extreme values
from <- min(forestinv$`2.5%`, na.rm=TRUE)-0.2
to <- max(forestinv$`97.5%`, na.rm=TRUE)+0.2
x1 <- seq(from = from, to = to, 0.01)
y1 <- dnorm(x1, mean = as.numeric(forestinv[1, "mean"]), sd = as.numeric(forestinv[1,"sd"]))
datam1 <- x1 %>% 
  cbind(y1) %>% 
  as_tibble() %>%
  rename(x = 1, 
         y = 2) %>% 
  mutate(Mechanism = "Propagule pressure")
x2 <- seq(from = from, to = to, 0.01)
y2 <- dnorm(x2, mean = as.numeric(forestinv[2, "mean"]), sd = as.numeric(forestinv[2,"sd"]))
datam2 <- x2 %>% 
  cbind(y2) %>% 
  as_tibble() %>%
  rename(x = 1, 
         y = 2) %>% 
  mutate(Mechanism = "Empty niches")
  
x3 <- seq(from = from, to = to, 0.01)
y3 <- dnorm(x3, mean = as.numeric(forestinv[3, "mean"]), sd = as.numeric(forestinv[3,"sd"]))
datam3inv <- x3 %>% 
  cbind(y3) %>% 
  as_tibble() %>%
  rename(x = 1, 
         y = 2) %>% 
  mutate(Mechanism = "Low biotic resistance") %>% 
  rbind(datam1, datam2) %>% mutate(type="invasive")
```

## Native community performance
### data preparation
```{r echo=TRUE}
# here, I merge the file with the estimated ES values with SD and calculate ES for the observations with no associated variability
BioticDataModel <- data %>% filter(ObsIDOriginal %in% NatUniqueBiotic$ObsIDOriginal) %>% 
  left_join(NatBioticEstimate, by = "ObsIDOriginal") %>% left_join(InvEstimate, by = "ObsIDOriginal") %>% 
  mutate(ESInvmean = ifelse(is.na(ESInvmean), ((InvTreatmean-InvContmean)/((InvTreatmean+InvContmean)/2)), ESInvmean),
         ESNatmean = ifelse(is.na(ESNatmean), ((NatTreatmean-NatContmean)/((NatTreatmean+NatContmean)/2)), ESNatmean),
         ObsID = row_number()) %>% group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()
```

### analysis
#### organizing model data
```{r}
J <- nrow(BioticDataModel) # number of observations
StudyID <- BioticDataModel$StudyID # data on study unique number (ID)
Mechanism <- BioticDataModel$Mechanism # data on mechanisms
natESavemean <- BioticDataModel$ESNatmean # data on ES mean
natESavesd <- BioticDataModel$ESInvSD # data on ES variability
M <- max(BioticDataModel$ESNatSD, na.rm = T) # maximum value of variability 
K <- length(unique(BioticDataModel$Mechanism)) # number of unique mechanisms
R <- K+1 # total number of tau
W <- length(unique(BioticDataModel$StudyID)) # number of unique study IDs

# data file to run model
datamodel <- list (J=J, StudyID=StudyID, Mechanism=Mechanism, 
                   natESavemean=natESavemean, natESavesd=natESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESNat <- function() {
  for(i in 1:J){
# missing values
natESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(natESavesd[i]*natESavesd[i])

natESavemean[i]~dnorm(iE[i],tau[Mechanism[i]])
natESavemean.h[i]~dnorm(iE[i],tau[Mechanism[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[Mechanism[i]]+iSRE[StudyID[i]]+ie[i]
residuals[i]<-natESavemean.h[i]-natESavemean[i]
}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ 
iSRE[i]~dnorm(0,tau[R])
}

  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}
} # end model

# initials 
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "natESavemean.h", "tau", "var", "residuals")

# running model
es.simNatDiffStudy <- bugs(datamodel, inits, parameters, model.file = modelESNat,
                           n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simNatDiffStudy[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simNatDiffStudy, "beta")
```

+ Model parameters
```{r parameters 2, echo=FALSE}
x<-as.data.frame(es.simNatDiffStudy[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESNat-Mech.csv"), row.names = F)
x
```

+ Model residuals
```{r residuals 2, include=FALSE}
resNat<-as.data.frame(es.simNatDiffStudy[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^residuals", rowname))

write.csv(resNat, file.path(parametersout, "resNat-MechRes.csv"), row.names = F)
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simNatDiffStudy[["summary"]]) 
natESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^natESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(natESavemean.h,natESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESNatmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x =-1, y = 0.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation for native community ES
samplesize <- BioticDataModel %>% 
  group_by(Mechanism) %>%
  add_tally() %>% ungroup() %>% select(Mechanism, n) %>% distinct() %>%
  arrange(Mechanism) %>%
  mutate(comparison2 = paste0("mechanism", 1:3))
         
nat <- as.data.frame(es.simNatDiffStudy[["summary"]])
nat <- nat %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(!grepl("^natESavemean.h", rowname),
         !grepl("^ESIA", rowname),
         !grepl("^tau", rowname),
         !grepl("^var", rowname),
         !grepl("^residuals",rowname),
         rowname != "deviance") %>%
  mutate(type = "native")
forestnat <- nat %>%
  arrange(rowname) %>%
  mutate(test = str_sub(rowname, 6, 6)) %>% 
  # mutate_all(na_if,"") %>% 
  mutate(comparison2 = ifelse(is.na(test), "overall", 
                             ifelse(!is.na(test), paste0('mechanism', test), NA))) %>%
  arrange(comparison2) %>%
  mutate(index = row_number()) %>%
  left_join(samplesize) %>%
  group_by(type) %>%
  mutate(n1 = ifelse(is.na(n), sum(n, na.rm=TRUE), n),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')'),
         index=ifelse(Mechanism==1, 3, ifelse(Mechanism==2, 1, 2))) %>%
  select(-n1, -test) %>%
  ungroup() 
```

```{r include=FALSE}
# adjust the x extreme values
from <- min(forestnat$`2.5%`, na.rm=TRUE)-0.2
to <- max(forestnat$`97.5%`, na.rm=TRUE)+0.2
x1 <- seq(from = from, to = to, 0.01)
y1 <- dnorm(x1, mean = as.numeric(forestnat[1, "mean"]), sd = as.numeric(forestnat[1,"sd"]))
datam1 <- x1 %>% 
  cbind(y1) %>% 
  as_tibble() %>%
  rename(x = 1, 
         y = 2) %>% 
  mutate(Mechanism = "Propagule pressure")
x2 <- seq(from = from, to = to, 0.01)
y2 <- dnorm(x2, mean = as.numeric(forestnat[2, "mean"]), sd = as.numeric(forestnat[2,"sd"]))
datam2 <- x2 %>% 
  cbind(y2) %>% 
  as_tibble() %>%
  rename(x = 1, 
         y = 2) %>% 
  mutate(Mechanism = "Empty niches")
  
x3 <- seq(from = from, to = to, 0.01)
y3 <- dnorm(x3, mean = as.numeric(forestnat[3, "mean"]), sd = as.numeric(forestnat[3,"sd"]))
datam3nat <- x3 %>% 
  cbind(y3) %>% 
  as_tibble() %>%
  rename(x = 1, 
         y = 2) %>% 
  mutate(Mechanism = "Low biotic resistance") %>% 
  rbind(datam1, datam2) %>% mutate(type="native")
```


```{r include=FALSE}
# forest plot
forestnat2 <- forestnat %>% mutate(treatment=ifelse(Mechanism==3,BRcolor,"white"),shape=23,index=index+0.2)
forestinv2 <- forestinv %>% arrange(Mechanism) %>% mutate(treatment=c(PPcolor,ENcolor,BRcolor),shape=25)

forest <- forestnat2 %>% rbind(forestinv2) %>% mutate(treatment2=ifelse(Mechanism==1, PPcolor,
                                                                     ifelse(Mechanism==2,ENcolor,BRcolor)),
                                                      index=ifelse(Mechanism==1,index-2,
                                                                   ifelse(Mechanism==2,index+2,index))) %>% arrange(index) 
ggplot(data=forest, 
       aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`)) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  geom_errorbarh(data=forest, height=.1, size=1,color=forest$treatment2)+
  geom_point(data=forest, size=4,shape=forest$shape,color=forest$treatment2,fill=forest$treatment,stroke=1)+ #, shape=15
  scale_y_continuous(name = "", breaks=1:nrow(forest), 
                     trans="reverse",
                     labels = c("Propagule pressure","Low biotic resistance", "Empty niches", "","","")) +
  theme_bw()+
  coord_cartesian(ylim = c(3.5,0.9), xlim=c(-1.25, 1.25), clip = "off")+
  xlab("Effect Size (mean+95%PI)")+
  theme(text=element_text(size=22, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title.x = element_text(size = 18),
      plot.margin = margin(1,1,1,1),#defines margin so the graph does not cut off 
      axis.line = element_line(colour = "black"),
      panel.border = element_blank() ) +    #remove top and right panel boundaries
  geom_text_repel(aes(label = n),direction="both",color=forest$treatment2,size=6,point.size = 6) +
  annotate("segment", x=-1.2, xend=1.2, y=3.55, yend=3.55, color="black", size=1.3, arrow=arrow(length = unit(0.03, "npc"), ends = 'last', type = "open")) +
  annotate("text", x=-1.15, y=3.45, label="lower", color="black", fontface = "italic", size=5) +
  annotate("text", x=1.15, y=3.45, label="higher", color="black", fontface = "italic", size=5)+
  annotate("text", x=0, y=3.45, label="Performance", color="black", fontface = "bold.italic", size=5)+
   annotate("text", x=-1.25, y=0.82, label="(a)", color="black", fontface = "bold", size=6)+
  annotate("text", x=-1.25, y=1.02, label="*", color="black", fontface = "bold", size=6)+
  annotate("text", x=-1.25, y=2.02, label="*", color="black", fontface = "bold", size=6)+
  annotate("text", x=-1.25, y=3.02, label="*", color="black", fontface = "bold", size=6)

# saves
ggsave(file.path(graph.wd,"ESNatInv_ForestPlot.png"), plot = last_plot(), width=6.8, height=5.2, dpi = 300, units="in")
```

```{r echo=FALSE}
# probability density plots
datam3 <- datam3nat %>% rbind(datam3inv)

p<- 
  ggplot(data = datam3, 
       mapping = aes(x=x, y=y, fill = as.factor(Mechanism), color=as.factor(Mechanism))) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  facet_grid(rows = vars(type),scales="free_y")+
  geom_line(aes(colour = as.factor(Mechanism), group = as.factor(Mechanism)), size=0.5) +
  geom_ribbon(aes(x = x, ymax = y, color=as.factor(Mechanism)), ymin=0, alpha=0.3) +
  scale_color_manual(values=c(ENcolor,BRcolor,PPcolor))+
    scale_fill_manual(values=c(ENcolor,BRcolor,PPcolor))+
  theme_bw()+
  xlab("Effect Size") +
  ylab("Probability Density Function") +
  scale_x_continuous(expand = c(0,0)) +
  theme(text=element_text(size=22, color="black"),
        legend.position="none",
        panel.grid = element_blank(),
        axis.title.x = element_text(size = 18),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.margin = margin(1,1,1,1))+#defines margin so the graph does not cut off ) +
  ggh4x::force_panelsizes(rows = unit(4.5/2, "in"), cols = unit(4, "in"))+
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)

egg::tag_facet(p, 
          x = -Inf, y = -Inf, 
          vjust = -12,	
          hjust = -0.1,
          size=6,
          open = "(", close = ")",
          tag_pool = c("b","c"))

# saves
ggsave(file.path(graph.wd,"ESNatInv_PDF.png"), plot = last_plot(), width=4.6, height=5.2, dpi = 300, units="in")
```

# **Q2: what are the abiotic conditions most suitable for invasion success?**

### data preparation
```{r}
AbioticDataModel <- data %>% 
  filter(ObsIDOriginal %in% NatUniqueAbiotic$ObsIDOriginal,) %>% 
  left_join(NatAbioticEstimate, by = "ObsIDOriginal") %>% left_join(InvEstimate, by = "ObsIDOriginal") %>% 
  # calculating the ES without SD
  mutate(ESInvmean = ifelse(is.na(ESInvSD), 
                            ((InvTreatmean-InvContmean)/((InvTreatmean+InvContmean)/2)), ESInvmean),
  ESNatmean = ifelse(is.na(ESNatSD), 
                                   ((NatTreatmean-NatContmean)/((NatTreatmean+NatContmean)/2)), ESNatmean)) %>%
  # fixing the directionality of the native ES (details on Text S1)
  mutate(ESNatmean = ifelse(TypeOfTreatment=="canopy_cover"| 
                              NativeResponse_SUBCATEGORY=="distance_from_river"|
           NativeResponse_SUBCATEGORY=="depth_soil_mottling"|
           NativeResponse_SUBCATEGORY=="distance_from_edge"|
           NativeResponse_SUBCATEGORY=="canopy_height_variability"|
           NativeResponse_SUBCATEGORY=="light_heterogeneity"|
           NativeResponse_SUBCATEGORY=="soil_bulk_density", -1*ESNatmean, ESNatmean)) %>% 
  # aggregating subcategories into fewer categories (details on Text S1)
  mutate(Abiotic_CATEGORY = ifelse(NativeResponse_SUBCATEGORY=="distance_from_river"|
                                       NativeResponse_SUBCATEGORY=="soil_moisture"|
                                       NativeResponse_SUBCATEGORY=="water_depth"|
                                       NativeResponse_SUBCATEGORY=="water_flow_pick"|
                                       NativeResponse_SUBCATEGORY=="water_flow_velocity"|
                                       NativeResponse_SUBCATEGORY=="water_surface_flow_permanence"|
                                       NativeResponse_SUBCATEGORY=="mean_water_potential"|
                                       NativeResponse_SUBCATEGORY=="depth_soil_mottling", "water_availability",
                                     ifelse(NativeResponse_SUBCATEGORY=="canopy_height_variability"|
                                       NativeResponse_SUBCATEGORY=="distance_from_edge"|
                                       NativeResponse_SUBCATEGORY=="light_availability"|
                                       NativeResponse_SUBCATEGORY=="light_heterogeneity", "light_availability",
                                       ifelse(NativeResponse_SUBCATEGORY=="CO2_concentration"|
                                       NativeResponse_SUBCATEGORY=="soil_nutrient_availability"|
                                       NativeResponse_SUBCATEGORY=="soil_pH"|
                                       NativeResponse_SUBCATEGORY=="standing_litter_biomass"|
                                       NativeResponse_SUBCATEGORY=="litter_biomass"|
                                       NativeResponse_SUBCATEGORY=="soil_bulk_density", "nutrient_availability", NA)))) %>% 
  filter(!is.na(Abiotic_CATEGORY)) %>% 
  group_by(StudyIDOriginal) %>%
  mutate(StudyID = cur_group_id()) %>% ungroup() %>% 
  group_by(Abiotic_CATEGORY) %>% 
  mutate(AbioticID = cur_group_id()) %>% 
  ungroup() %>% 
  mutate(ObsID = row_number())
```

### analysis
#### organizing the data
```{r}
J <- nrow(AbioticDataModel)
StudyID <- AbioticDataModel$StudyID
Category <- AbioticDataModel$AbioticID
invESavemean <- AbioticDataModel$ESInvmean
invESavesd <- AbioticDataModel$ESInvSD
natESavemean <- AbioticDataModel$ESNatmean
natESavesd <- AbioticDataModel$ESNatSD
C <- length(unique(AbioticDataModel$AbioticID))
M <- max(AbioticDataModel$ESInvSD, na.rm = T) 
N <- max(AbioticDataModel$ESNatSD, na.rm = T) 
K <- length(unique(AbioticDataModel$AbioticID))
Q <- K-2
G <- K-1
R <- K+1
H <- nrow(AbioticDataModel[AbioticDataModel$AbioticID == 1,]) # when light availability
L <- nrow(AbioticDataModel[AbioticDataModel$AbioticID == 2,]) # when nutrient availability
P <- nrow(AbioticDataModel[AbioticDataModel$AbioticID == 3,]) # when water availability
O <- length(seq(-2,2,0.1))
natpred<-seq(-2,2,0.1)
W <- length(unique(AbioticDataModel$StudyID))

datamodel <- list (J=J, Category=Category, invESavemean=invESavemean, invESavesd=invESavesd, 
                   natESavemean=natESavemean, natESavesd=natESavesd, C=C, M=M, N=N, K=K, G=G, 
                   H=H, L=L, O=O, natpred=natpred, W=W, StudyID=StudyID, R=R,Q=Q, G=G, P=P)
```

#### model
```{r}
modelESInvAbioticPred <- function() {

for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,)
natESavesd[i]~dnorm(N,1)%_%C(0,)

tauI[i]<-1/(invESavesd[i]*invESavesd[i])
tauN[i]<-1/(natESavesd[i]*natESavesd[i])

invESavemean[i]~dnorm(E[i],tau[Category[i]])
invESavemean.h[i]~dnorm(E[i],tau[Category[i]]) 
natES[i]~dnorm(natESavemean[i],tauN[i])
e[i]~dnorm(0,tauI[i])
E[i]<-alpha[Category[i]]+beta[Category[i]]*natES[i]+iSRE[StudyID[i]]+e[i]
residuals[i]<-invESavemean.h[i]-invESavemean[i]
}

# priors
for(i in 1:C){ 
alpha[i]~dnorm(0,0.0001)
beta[i]~dnorm(0,0.0001)
}

for(i in 1:W){
  iSRE[i]~dnorm(0,tau[R])
}
  
for(i in 1:R){
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

# predictions
taup[Q]<-H*tau[Q] 
taup[G]<-L*tau[G] 
taup[K]<-P*tau[K]
taup[R]<-W*tau[R]

studyRE~dnorm(0,taup[R])

  for(m in 1:C){ 
error[m]~dnorm(0,taup[m])
for(j in 1:O){
invpred[m,j]<-alpha[m]+beta[m]*natpred[j]+studyRE+error[m]
  }
}

}# end model

# parameters
parameters <- c("alpha", "beta", "invESavemean.h", "invpred", "tau", "var", "residuals")

# initials 
inits <- list(
  list(alpha = c(rep(0, times=K)), beta = c(rep(0, times=K)), var = c(rep(1, times=R))),
  list(alpha = c(rep(0, times=K)), beta = c(rep(1, times=K)), var = c(rep(1, times=R))),
  list(alpha = c(rep(0, times=K)), beta = c(rep(-1, times=K)), var = c(rep(1, times=R)))
)

# running model
es.simAbioticPred <- bugs(datamodel, inits, parameters, model.file = modelESInvAbioticPred, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value: 
```{r echo=FALSE}
es.simAbioticPred[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simAbioticPred, "beta")
```

+ Model parameters
```{r parameters 3, echo=FALSE}
x<-as.data.frame(es.simAbioticPred[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname)|
                                                     grepl("^alpha", rowname))
write.csv(x, file.path(parametersout, "ESInv~resourceAvailability.csv"), row.names = F)
x
```

+ Model residuals
```{r residuals 3, include=FALSE}
resInvAbiotic<-as.data.frame(es.simAbioticPred[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^residuals", rowname))

write.csv(resInvAbiotic, file.path(parametersout, "ESInv~resourceAvailabilityRes.csv"), row.names = F)
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simAbioticPred[["summary"]]) 
invESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^invESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(invESavemean.h,invESavemean)
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESInvmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.5, y = 1, labels = mylabel)
```

```{r include=FALSE}
# data preparation for forest plot (slope)
samplesize <- AbioticDataModel %>% group_by(Abiotic_CATEGORY, AbioticID) %>% tally() %>% 
   mutate(n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
         
forest <- as.data.frame(es.simAbioticPred[["summary"]])
forest <- forest %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  arrange(rowname) %>%
  mutate(AbioticID = seq(1:3),
         type = "beta") %>%
  mutate(index = row_number()) %>%
  left_join(samplesize) %>% 
  mutate(index=AbioticID)
```

```{r include=FALSE}
# x-axis label
xname <- "Slope (mean+95%CI)"

#forest plot of slope
ggplot(data=forest, aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`))+ 
   geom_errorbarh(data=forest, height=.1, size=1,aes(color=Abiotic_CATEGORY)) +
  geom_point(data = forest, size=5, color=c(Lightcolor,Nutrientcolor,Watercolor),shape=c(21,24,22),fill="white",stroke=1)+
 
  scale_x_continuous(limits=c(-0.5,0.5), name=xname) +
scale_y_continuous(name = "", breaks=1:nrow(forest),trans="reverse", labels = c("Light\navailability","Nutrient\navailability","Water\navailability")) +
  scale_color_manual(values=c(Lightcolor,Nutrientcolor,Watercolor)) +
geom_vline(xintercept=0, linetype="dashed", alpha=.5, size=1)+
theme_bw()+
  labs(y=NULL)+
  coord_cartesian(ylim = c(3,1), xlim=c(-.5, .5), clip = "off")+
theme(text=element_text(size=22, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title.x = element_text(size = 18, color="black"),
      plot.margin = margin(1,1,1,10),
      panel.background = element_rect(fill='transparent', color=NA),
      plot.background = element_rect(fill='transparent', color=NA),
      axis.text.y=element_blank(),  #remove y axis labels
      axis.ticks.y=element_blank(),
      axis.title.y=element_blank(),
      axis.text.x = element_text(size=19),
      axis.line = element_line(colour = "black"),
      panel.border = element_blank())+ 
  geom_text_repel(aes(label = n),direction="both",color=c(Lightcolor,Nutrientcolor,Watercolor),size=6, point.size=12,segment.colour = NA)+
   annotate("text", x=-0.5, y=0.95, label="(b)", color="black", fontface = "bold", size=6)

# saves
ggsave(file.path(graph.wd,"ESInvAbiotic_PredSlope.png"), plot = last_plot(), width=5, height=6, dpi = 300, units="in")
```

```{r model1: ESInv~ESNat EN data prep, include=FALSE}
## data preparation to plot predictions
ESInvPred<-as.data.frame(es.simAbioticPred[["summary"]])
ESInvPred <- ESInvPred %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^invpred", rowname)) %>% 
  mutate(ESNatPred = rep(seq(-2,2,0.1), times = 3),
         AbioticID = rep(1:3, times=1, each =41)) %>% 
  left_join(select(AbioticDataModel, c(Abiotic_CATEGORY, AbioticID)), by = "AbioticID") %>%
  distinct() %>% 
  rename(ESInvPred = mean,
         CI25 = `2.5%`,
         CI975 = `97.5%`)
```

```{r model1: ESInv~ESNat EN plot, include=FALSE}
# plotting real data in the background and the prediction lines on top
  ggplot(ESInvPred, aes(ESNatPred)) +
  geom_hline(yintercept=0, linetype="dashed",
             color = "gray", size=1.5) +
  geom_line(aes(y=ESInvPred, colour = as.factor(Abiotic_CATEGORY), group = as.factor(Abiotic_CATEGORY)), size=1.5) +
  geom_point(data = AbioticDataModel, aes(ESNatmean, ESInvmean, 
                                     color = as.factor(Abiotic_CATEGORY),
                                     group = as.factor(Abiotic_CATEGORY),
                                     shape = as.factor(Abiotic_CATEGORY)),size=3,alpha=0.8) +
  theme_bw()+
  scale_x_continuous(expand = c(0,0)) +
  coord_cartesian(ylim = c(-0.2, 2.1), 
                  xlim=c(-2,2), 
                  clip = "off")+
  geom_ribbon(aes(ymin=CI25, ymax=CI975, group = as.factor(Abiotic_CATEGORY), fill = as.factor(Abiotic_CATEGORY)), alpha=0.2) +
  scale_color_manual(values=c(Lightcolor,Nutrientcolor,Watercolor),
                     labels=c("Light","Soil nutrient","Soil water")) +
  scale_fill_manual(values=c(Lightcolor,Nutrientcolor,Watercolor), 
                    labels=c("Light","Soil nutrient","Soil water")) +
  scale_shape_manual(values=c(16,17,15),labels=c("Light","Soil nutrient","Soil water")) +
  xlab("Abiotic Effect Size") +
  ylab("Invasive Effect Size (mean+95%PI)\n\n") +
  theme(text=element_text(size=20, color="black"),
        legend.position="bottom",
        legend.margin=margin(t=-10), # removes space between graph and legend
        legend.title = element_blank(),
        panel.grid = element_blank(),
        legend.text = element_text(size = 17),
        legend.key.size = unit(0.4, "cm"),
      plot.margin = margin(1,10,1,1),#defines margin so the graph does not cut off  +
      axis.line = element_line(colour = "black"),
      axis.text.y = element_text(size=19),
      axis.text.x = element_text(size=19),
      panel.border = element_blank())+
  annotate("segment", x=-1.9, xend=1.9, y=-0.25, yend=-0.25, color="black", size=1.3, arrow=arrow(length = unit(0.03, "npc"), ends = 'last', type = "open")) +
  annotate("text", x=-1.8, y=-0.15, label="lower", color="black", fontface = "italic",size=6) +
  annotate("text", x=1.8, y=-0.15, label="higher", color="black", fontface = "italic",size=6)+
  annotate("text", x=0, y=-0.15, label="Resource availability", color="black", fontface = "bold.italic",size=6) +
  annotate("segment", y=0.05, yend=1.95, x=-2.35, xend=-2.35, color="black", size=1.3, arrow=arrow(length = unit(0.03, "npc"), ends = 'last', type = "open")) +
  annotate("text", y=0.05, x=-2.52, label="lower", color="black", fontface = "italic",angle=90, size=6) +
  annotate("text", y=1.95, x=-2.52, label="higher", color="black", fontface = "italic", angle=90, size=6)+
  annotate("text", y=1, x=-2.52, label="Invader performance", color="black", fontface = "bold.italic", angle=90, size=6)+
   annotate("text", x=-1.87, y=2.15, label="(a)", color="black", fontface = "bold", size=6)

# saves
ggsave(file.path(graph.wd,"ESInvAbiotic_Pred.png"), plot = last_plot(), width=8, height=6, dpi = 300, units="in")
```

# **Q3: Through which mechanism of invasion, the native community is most vulnerable?**
### data preparation
```{r echo=TRUE}
BioticDataModel <- data %>% filter(ObsIDOriginal %in% NatUniqueBiotic$ObsIDOriginal) %>% 
  left_join(NatBioticEstimate, by = "ObsIDOriginal") %>% left_join(InvEstimate, by = "ObsIDOriginal") %>% 
  mutate(ESInvmean = ifelse(is.na(ESInvmean), ((InvTreatmean-InvContmean)/((InvTreatmean+InvContmean)/2)), ESInvmean),
         ESNatmean = ifelse(is.na(ESNatmean), ((NatTreatmean-NatContmean)/((NatTreatmean+NatContmean)/2)), ESNatmean),
         ObsID = row_number()) %>% group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()
```

### organizing model data
```{r}
# observations per mechanism
obs <- BioticDataModel %>% count(Mechanism)

J <- nrow(BioticDataModel)
StudyID <- BioticDataModel$StudyID
Mechanism <- BioticDataModel$Mechanism
invESavemean <- BioticDataModel$ESInvmean
invESavesd <- BioticDataModel$ESInvSD
natESavemean <- BioticDataModel$ESNatmean
natESavesd <- BioticDataModel$ESNatSD
W <- length(unique(BioticDataModel$StudyID))
M <- max(BioticDataModel$ESInvSD, na.rm = T)
N <- max(BioticDataModel$ESNatSD, na.rm = T)
invpred<-seq(0,2,0.1)
K <- length(unique(BioticDataModel$Mechanism))
R <- K+1
U <- as.numeric(obs[1,2]) #mechanism1
V <- as.numeric(obs[2,2]) #mechanism2
X <- as.numeric(obs[3,2]) #mechanism3
O <- length(seq(0,2,0.1))


datamodel <- list (J = J, StudyID = StudyID, Mechanism = Mechanism, invESavemean = invESavemean, invESavesd = invESavesd, natESavemean = natESavemean, natESavesd = natESavesd, W=W, invpred = invpred, M=M, N=N, K=K, R=R, U=U, V=V, X=X, O=O)
```

### model
```{r}
modelBiotic1 <- function() {

for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,)
natESavesd[i]~dnorm(N,1)%_%C(0,)

tauI[i]<-1/(invESavesd[i]*invESavesd[i])
tauN[i]<-1/(natESavesd[i]*natESavesd[i])

natESavemean[i]~dnorm(E[i],tau[Mechanism[i]])
natESavemean.h[i]~dnorm(E[i],tau[Mechanism[i]]) 
invES[i]~dnorm(invESavemean[i],tauI[i])
e[i]~dnorm(0,tauN[i])
E[i]<-beta[Mechanism[i]]*invES[i]+iSRE[StudyID[i]]+e[i] # intercept == 0
residuals[i]<-natESavemean.h-natESavemean
}

# priors
for(i in 1:K){
beta[i]~dnorm(0,0.0001)
}

for(i in 1:W){ # SRE
iSRE[i]~dnorm(0,tau[R])
}

for(i in 1:R){
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

# predictions
  
taup[1]<-U*tau[1] #mechanism1
taup[2]<-V*tau[2] #mechanism2
taup[3]<-X*tau[3] #mechanism3
taup[4]<-W*tau[4] #SRE

studyRE~dnorm(0,taup[R])

  for(m in 1:K){
error[m]~dnorm(0,taup[m])
study[m]~dnorm(0,taup[m])
for(j in 1:O){
natpred[m,j]<-beta[m]*invpred[j]+studyRE+error[m]
  }
}


}# end model

# parameters
parameters <- c("beta", "natESavemean.h", "natpred", "tau", "var", "residuals")

# initials
inits <- list(
  list(beta = c(rep(0, times=K)), var = c(rep(1, times=R))),
  list(beta = c(rep(1, times=K)), var = c(rep(1, times=R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times=R)))
)

# running model
es.simBiotic <- bugs(datamodel, inits, parameters, model.file = modelBiotic1, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simBiotic[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simBiotic, "beta")
```

+ Model parameters
```{r parameters 4, echo=FALSE}
x<-as.data.frame(es.simBiotic[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESNat-impactMech.csv"), row.names = F)
x
```

+ Model residuals
```{r residuals 4, include=FALSE}
resNatImpact<-as.data.frame(es.simBiotic[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^residuals", rowname))

write.csv(resNatImpact, file.path(parametersout, "ESNat-impactMechRes.csv"), row.names = F)
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simBiotic[["summary"]]) 
natESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^natESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(natESavemean.h,natESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESNatmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = -1.7, y = 0, labels = mylabel)
```

```{r include=FALSE}
# organizing data for forest plot (slopes, inset)
samplesize <- BioticDataModel %>% group_by(Mechanism) %>% add_tally() %>% ungroup() %>% select(Mechanism, n) %>% distinct() %>%
  arrange(Mechanism) %>% mutate(n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
         
forest <- as.data.frame(es.simBiotic[["summary"]])
forest <- forest %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(!grepl("^natESavemean.h", rowname),!grepl("^B.h", rowname), rowname != "deviance", !grepl("^alpha", rowname), !grepl("^natpred", rowname),!grepl("^tau", rowname),!grepl("^var", rowname),!grepl("^residuals", rowname)) %>%
  arrange(rowname) %>%
  mutate(Mechanism = seq(1:3),
         type = "beta") %>%
  mutate(index = row_number()) %>%
  left_join(samplesize) %>% 
  mutate(Mechanismdummy = ifelse(index==1, "Propagule pressure", ifelse(index==2, "Empty niches", "Low biotic resistance")),
  index=ifelse(Mechanism==1, 3, ifelse(Mechanism==2, 1, 2)),
  shape=ifelse(Mechanism==1, 22, ifelse(Mechanism==2, 21, 24)),
  treatment2=ifelse(Mechanism==1, PPcolor, ifelse(Mechanism==2, ENcolor, BRcolor)),
  treatment=ifelse(Mechanism==3, BRcolor, "white"))
```

```{r include=FALSE}
# forest plot (slopes, inset)

# x-axis label
xname <- "Slope (mean+95%CI)"

inset_plot <-
  ggplot(data=forest, aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`))+ 
  geom_errorbarh(data=forest, height=.1, size=1,color=forest$treatment2) +
  geom_point(data = forest, size=4, shape=forest$shape, color=forest$treatment2,fill=forest$treatment,stroke=1)+
    scale_x_continuous(limits=c(-0.8,0.15), name=xname, breaks = c(-0.8, -0.4, 0)) +
  scale_y_continuous(name = "", breaks=1:nrow(forest),trans="reverse", labels=NULL) +
  geom_vline(xintercept=0, linetype="dashed", alpha=.5, size=1)+
  theme_bw()+
  coord_cartesian(ylim=c(3.1,0.9),clip = "off")+
  theme(panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.text.x=element_text(size = 13),
      axis.title = element_text(size = 13, color="black"),
      panel.background = element_rect(fill='transparent', color=NA),
      plot.background = element_rect(fill='transparent', color=NA),
      axis.ticks.y=element_blank(),
      axis.title.y=element_blank())+ 
  annotate("text", x=-0.45, y=2.5, label="(15)", color=PPcolor, size=5)+
  annotate("text", x=-0.45, y=1.5, label="(52)", color=BRcolor, size=5)+
  annotate("text", x=-0.09, y=1.4, label="(36)", color=ENcolor, size=5)
``` 


```{r include=FALSE}
# data prep for: ESNat~ESInv with predictions by mechanism + inset

## Real data (calculated values)
BioticDataModel<- BioticDataModel %>% 
   mutate(Mechanismdummy = ifelse(Mechanism ==1, "Propagule pressure", ifelse(Mechanism==2, "Empty niches", "Low biotic resistance")),
  shape=ifelse(Mechanism==1, 19, ifelse(Mechanism==2, 17, 15)))
## predictions
ESNatPred<-as.data.frame(es.simBiotic[["summary"]])
ESNatPred <- ESNatPred %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^natpred", rowname)) %>% 
  mutate(ESInvPred = rep(seq(0,2,0.1), times = 3),
         Mechanism = rep(1:3, times=1, each =21),
         Mechanismdummy = ifelse(Mechanism ==1, "Propagule pressure",
                              ifelse(Mechanism==2, "Empty niches", "Low biotic resistance"))) %>% 
  rename(ESNatPred = mean,
         CI25 = `2.5%`,
         CI975 = `97.5%`)
```

```{r model1: ESNat~ESInv plot, include=FALSE}
## ESNat~ESInv with predictions by mechanism + inset
main_plot <-
  ggplot(ESNatPred, aes(ESInvPred)) +
  geom_hline(yintercept=0, linetype="dashed",
             color = "gray", size=1.5) +
  geom_point(data = BioticDataModel, aes(invESavemean, natESavemean,
                                     color = as.factor(Mechanismdummy),
                                     group = as.factor(Mechanismdummy),
                                     shape=as.factor(Mechanismdummy)),size=3,alpha=0.8) +
  geom_line(aes(y=ESNatPred, colour = as.factor(Mechanismdummy), group = as.factor(Mechanismdummy)), size=1.5) +
  theme_bw()+
  scale_x_continuous(expand = c(0,0)) +
  coord_cartesian(ylim = c(-2.1, 2), xlim=c(0,2), clip = "off")+
  scale_color_manual(values=c(ENcolor,BRcolor,PPcolor))+
    scale_fill_manual(values=c(ENcolor,BRcolor,PPcolor))+
  geom_ribbon(aes(ymin=CI25, ymax=CI975, group = as.factor(Mechanismdummy), fill = as.factor(Mechanismdummy)), alpha=0.2) +
  xlab("Invasive Effect Size") +
  ylab("Impact on native community\n\n ") +
  guides(color = guide_legend(nrow = 1))+
  theme(legend.position=c(0.4,-0.17),
        legend.margin=margin(t=-10),
        legend.title = element_blank(),
        panel.grid = element_blank(),
        legend.text = element_text(size = 15),
        legend.key.size = unit(0.4, "cm"),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 22),
        axis.text.y = element_text(size=19),
      axis.text.x = element_text(size=19),
        axis.line = element_line(colour = "black"),
      panel.border = element_blank(),
      plot.margin = margin(1,10,18,1))+  #defines margin so the graph does not cut off) 
  annotate("segment", x=0.1, xend=1.9, y=-2.16, yend=-2.16, color="black", size=1.3, arrow=arrow(length = unit(0.03, "npc"), ends = 'last', type = "open")) +
  annotate("text", x=0.2, y=-1.95, label="lower", color="black", fontface = "italic", size=6) +
  annotate("text", x=1.8, y=-1.95, label="higher", color="black", fontface = "italic", size=6)+
  annotate("text", x=1, y=-1.95, label="Performance", color="black", fontface = "bold.italic", size=6) +
  annotate("segment", y=2, yend=-1.9, x=-0.15, xend=-0.15, color="black", size=1.3, arrow=arrow(length = unit(0.03, "npc"), ends = 'last', type = "open")) +
  annotate("text", y=-1.9, x=-0.23, label="higher", color="black", fontface = "italic",angle=90, size=6) +
  annotate("text", y=1.9, x=-0.23, label="lower", color="black", fontface = "italic", angle=90, size=6)+
  annotate("text", y=0, x=-0.23, label="Impact", color="black", fontface = "bold.italic", angle=90, size=6)+
   annotate("text", x=0.08, y=2.14, label="(a)", color="black", fontface = "bold", size=6)+
  annotate("text", y=0, x=-0.325, label="Effect Size (mean+95%PI)", color="black", angle=90,size=6)

# adds the inset graph
main_plot +
patchwork::inset_element(inset_plot,0.74, 0.66,0.98, 0.999, align_to = 'plot') # overlaps graph to create inset

# saves
ggsave(file.path(graph.wd,"ESNatInv_Predictions.png"), plot = last_plot(), width=7, height=6, dpi = 300, units="in")
```

# **Q4: Is the native community more impacted by one of multiple invaders?**
### organizing model data
```{r}
#observations per mechanism
BioticDataModel <- BioticDataModel %>% mutate(MultipleInvader=ifelse(MultipleInvader==0, 2, 1))

obs <- BioticDataModel %>% count(MultipleInvader)

J <- nrow(BioticDataModel)
StudyID <- BioticDataModel$StudyID
Category <- BioticDataModel$MultipleInvader
invESavemean <- BioticDataModel$ESInvmean
invESavesd <- BioticDataModel$ESInvSD
natESavemean <- BioticDataModel$ESNatmean
natESavesd <- BioticDataModel$ESNatSD
W <- length(unique(BioticDataModel$StudyID))
M <- max(BioticDataModel$ESInvSD, na.rm = T)
N <- max(BioticDataModel$ESNatSD, na.rm = T)
invpred<-seq(0,2,0.1)
K <- length(unique(BioticDataModel$MultipleInvader))
R <- K+1
U <- as.numeric(obs[1,2]) #multiple invader
V <- as.numeric(obs[2,2]) #single invader
O <- length(seq(0,2,0.1))

datamodel <- list (J = J, StudyID = StudyID, Category = Category, invESavemean = invESavemean, invESavesd = invESavesd, natESavemean = natESavemean, natESavesd = natESavesd, W=W, invpred = invpred, M=M, N=N, K=K, R=R, U=U, V=V, O=O)
```

### model
```{r}
modelBiotic2 <- function() {

for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,)
natESavesd[i]~dnorm(N,1)%_%C(0,)

tauI[i]<-1/(invESavesd[i]*invESavesd[i])
tauN[i]<-1/(natESavesd[i]*natESavesd[i])

natESavemean[i]~dnorm(E[i],tau[Category[i]])
natESavemean.h[i]~dnorm(E[i],tau[Category[i]]) #predicted
invES[i]~dnorm(invESavemean[i],tauI[i])
e[i]~dnorm(0,tauN[i])
E[i]<-beta[Category[i]]*invES[i]+iSRE[StudyID[i]]+e[i] # intercept == 0

}

# priors
for(i in 1:K){
beta[i]~dnorm(0,0.0001)
}

for(i in 1:W){ # SRE
iSRE[i]~dnorm(0,tau[R])
}

for(i in 1:R){
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

# predictions
  
taup[1]<-U*tau[1] #multiple invader
taup[2]<-V*tau[2] #single invader
taup[3]<-W*tau[3] #SRE

studyRE~dnorm(0,taup[R])

  for(m in 1:K){
error[m]~dnorm(0,taup[m])
study[m]~dnorm(0,taup[m])
for(j in 1:O){
natpred[m,j]<-beta[m]*invpred[j]+studyRE+error[m]
  }
}


}# end model

# parameters
parameters <- c("beta", "natESavemean.h", "natpred", "tau", "var", "residuals")

# initials
inits <- list(
  list(beta = c(rep(0, times=K)), var = c(rep(1, times=R))),
  list(beta = c(rep(1, times=K)), var = c(rep(1, times=R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times=R)))
)

# running model
es.simBiotic2 <- bugs(datamodel, inits, parameters, model.file = modelBiotic2, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simBiotic2[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simBiotic2, "beta")
```

+ Model parameters
```{r parameters 5, echo=FALSE}
x<-as.data.frame(es.simBiotic2[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESNat~impactMultInv.csv"), row.names = F)
x
```

+ Model residuals
```{r residuals 5, include=FALSE}
resNatMultInv<-as.data.frame(es.simBiotic2[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^residuals", rowname))

write.csv(resNatMultInv, file.path(parametersout, "ESNat~impactMultInvRes.csv"), row.names = F)
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simBiotic2[["summary"]])
natESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^natESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(natESavemean.h,natESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESNatmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = -1.7, y = 0, labels = mylabel)
```

```{r include=FALSE}
# organizing data for forest plot (slopes, inset)
samplesize <- BioticDataModel %>% group_by(MultipleInvader) %>% add_tally() %>% ungroup() %>% select(MultipleInvader, n) %>% distinct() %>%
  arrange(MultipleInvader) %>% mutate(comparison2 = paste0("multipleinvader", 1:0), 
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 

forest <- as.data.frame(es.simBiotic2[["summary"]])
forest <- forest %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(!grepl("^natESavemean.h", rowname),!grepl("^B.h", rowname), rowname != "deviance", !grepl("^alpha", rowname), !grepl("^natpred", rowname),!grepl("^tau", rowname), !grepl("^var", rowname), !grepl("^residuals", rowname)) %>%
  arrange(rowname) %>%
  mutate(comparison2 = paste0("multipleinvader", 1:0),
         type = "beta") %>%
  mutate(index = row_number()) %>%
  left_join(samplesize) %>% 
  mutate(MultInvdummy = ifelse(index==1, "Multiple invaders", "Single invader"),
  index=ifelse(comparison2=="multipleinvader1", 2, 1))
```

```{r model1: Slope, include=FALSE}
# forest plot (slopes, inset)

xname <- "Slope (mean+95%CI)"
inset_plot <-
  ggplot(data=forest, aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`))+ 
  geom_errorbarh(data=forest, height=.1, size=1,color=c(Multicolor,Singlecolor)) +
  geom_point(data = forest, size=4, shape=c(24,21), stroke=1,color=c(Multicolor,Singlecolor),fill=c("white",Singlecolor))+
    scale_x_continuous(limits=c(-0.6,.5), name=xname) +
scale_y_continuous(name = "", breaks=1:nrow(forest),trans="reverse", labels = c("Single invader", "Multiple Invader")) +
geom_vline(xintercept=0, linetype="dashed", alpha=.5, size=1)+
theme_bw()+
theme(panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title = element_text(size = 13, color="black"),
      axis.text.x=element_text(size = 13),
      panel.background = element_rect(fill='transparent', color=NA),
      plot.background = element_rect(fill='transparent', color=NA),
      axis.text.y=element_blank(),  #remove y axis labels
      axis.ticks.y=element_blank(),
      axis.title.y=element_blank())+ 
  annotate("text", x=-0.4, y=1.2, label="(77)", color=Singlecolor, size=5)+
   annotate("text", x=0.2, y=1.75, label="(26)", color=Multicolor, size=5)
``` 

```{r include=FALSE}
# data prep for ESNat~ESInv with predictions by number of invaders + inset
## Real data (calculated values)
BioticDataModel<- BioticDataModel %>% 
   mutate(MultInvdummy = ifelse(MultipleInvader==1, "Multiple invaders", "Single invader"))

range(BioticDataModel$ESInvmean)
range(BioticDataModel$ESNatmean)

## predictions
ESNatPred<-as.data.frame(es.simBiotic2[["summary"]])
ESNatPred <- ESNatPred %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^natpred", rowname)) %>% 
  mutate(ESInvPred = rep(seq(0,2,0.1), times = 2),
         MultipleInvader = rep(1:0, times=1, each =21),
         MultInvdummy = ifelse(MultipleInvader==1, "Multiple invaders", "Single invader")) %>% 
  rename(ESNatPred = mean,
         CI25 = `2.5%`,
         CI975 = `97.5%`)
```

```{r echo=FALSE}
# ESNat~ESInv with predictions by number of invaders + inset

main_plot <-
  ggplot(ESNatPred, aes(ESInvPred)) +
  geom_hline(yintercept=0, linetype="dashed",
             color = "gray", size=1.5) +
  geom_point(data = BioticDataModel, aes(invESavemean, natESavemean,
                                     color = as.factor(MultInvdummy),
                                     group = as.factor(MultInvdummy),
                                     shape = as.factor(MultInvdummy)),size=3,alpha=0.8) +
  geom_line(aes(y=ESNatPred, colour = as.factor(MultInvdummy), group = as.factor(MultInvdummy)), size=1.5) +
  theme_bw()+
  scale_x_continuous(expand = c(0,0)) +
  coord_cartesian(ylim = c(-2.1, 2), xlim=c(0,2), clip = "off")+
  scale_color_manual(values=c(Multicolor,Singlecolor))+
    scale_fill_manual(values=c(Multicolor,Singlecolor))+
  geom_ribbon(aes(ymin=CI25, ymax=CI975, group = as.factor(MultInvdummy), fill = as.factor(MultInvdummy)), alpha=0.2) +
  xlab("Invasive Effect Size") +
  ylab("\n\n") +
  guides(color = guide_legend(nrow = 1))+
  theme(legend.position=c(0.5,-0.17),
        legend.margin=margin(t=-10),
        legend.title = element_blank(),
        panel.grid = element_blank(),
        legend.text = element_text(size = 15),
        legend.key.size = unit(0.4, "cm"),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 22),
        axis.text.y = element_text(size=19),
        axis.text.x = element_text(size=19),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        plot.margin = margin(1,10,18,1))+  #defines margin so the graph does not cut off) 
  annotate("segment", x=0.1, xend=1.9, y=-2.16, yend=-2.16, color="black", size=1.3, arrow=arrow(length = unit(0.03, "npc"), ends = 'last', type = "open")) +
  annotate("text", x=0.2, y=-2, label="lower", color="black", fontface = "italic", size=6) +
  annotate("text", x=1.8, y=-2, label="higher", color="black", fontface = "italic", size=6)+
  annotate("text", x=1, y=-2, label="Performance", color="black", fontface = "bold.italic", size=6) +
   annotate("text", x=0.08, y=2.15, label="(b)", color="black", fontface = "bold", size=6)#+

# adding the inset graph
main_plot +
patchwork::inset_element(inset_plot,0.74, 0.66,0.98, 0.999, align_to = 'plot') # overlaps graph to create inset

ggsave(file.path(graph.wd,"ESNatInv_MultInvPredictions.png"), plot = last_plot(), width=7, height=6, dpi = 300, units="in")
```

# **SM: Model residuals per covariates**
```{r}
resInv
resNat
resInvAbiotic
resNatImpact
resNatMultInv
```

## Ecoregions
## Forest community
## Forest type
## Study type
## Disturbance
## Type of distubance

# **SM: Invasive species and native community Performance per variables**
## Ecoregions
### Invasive performance
#### organizing model data
```{r}
InvAllDataModel2 <- InvAllDataModel %>% filter(!is.na(OlsonEtAl2001_GlobalEcoregions))%>% 
  group_by(OlsonEtAl2001_GlobalEcoregions) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(InvAllDataModel2)
StudyID <- InvAllDataModel2$StudyID
CategoryID <- InvAllDataModel2$CategoryID
invESavemean <- InvAllDataModel2$ESInvmean
invESavesd <- InvAllDataModel2$ESInvSD
M <- max(InvAllDataModel2$ESInvSD, na.rm = T) 
K <- length(unique(InvAllDataModel2$CategoryID))
R <- K+1 
W <- length(unique(InvAllDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, invESavemean=invESavemean, invESavesd=invESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(invESavesd[i]*invESavesd[i])

invESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
invESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ # SRE
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "invESavemean.h", "tau", "var") 

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 8, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESInv-Ecoregions.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) 
invESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^invESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(invESavemean.h,invESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESInvmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 1.5, labels = mylabel)
```

```{r include=FALSE}
# organizing data to plot it later together with native community ES
samplesize <- InvAllDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, OlsonEtAl2001_GlobalEcoregions, n) %>% distinct() 

         
inv <- as.data.frame(es.simCat[["summary"]])
inv <- inv %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "invasive")

forest2inv <- inv %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
  
```

### Native performance
#### organizing model data
```{r}
BioticDataModel2 <- BioticDataModel %>% filter(!is.na(OlsonEtAl2001_GlobalEcoregions))%>% 
  group_by(OlsonEtAl2001_GlobalEcoregions) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(BioticDataModel2)
StudyID <- BioticDataModel2$StudyID
CategoryID <- BioticDataModel2$CategoryID
natESavemean <- BioticDataModel2$ESNatmean
natESavesd <- BioticDataModel2$ESNatSD
M <- max(BioticDataModel2$ESNatSD, na.rm = T) 
K <- length(unique(BioticDataModel2$CategoryID))
R <- K+1 
W <- length(unique(BioticDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, natESavemean=natESavemean, natESavesd=natESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
natESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(natESavesd[i]*natESavesd[i])

natESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
natESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ # SRE
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials 
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "natESavemean.h", "tau", "var")

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 14, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESNat-Ecoregions.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]])
natESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^natESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(natESavemean.h,natESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESNatmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 0.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- BioticDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, OlsonEtAl2001_GlobalEcoregions, n) %>% distinct() 
         
nat <- as.data.frame(es.simCat[["summary"]])
nat <- nat %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "native")

forest2nat <- nat %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
  
```

```{r include=FALSE}
# forest plot of ecoregions, with invasive species and native community ES
forest2nat <- forest2nat %>% mutate(index=ifelse(CategoryID==1, 2.2,
                                       ifelse(CategoryID==2, 5.2,
                                       ifelse(CategoryID>2, index+3.2, index))),
                                 treatment=ifelse(CategoryID==3, "a", "b"),
                                 treatment2=ESNatcolor,shape=23)

forest2 <-  forest2inv %>%  mutate(treatment=ifelse(CategoryID==1|CategoryID==3|CategoryID==5|
                                                        CategoryID==9, "d", "c"), 
                                   treatment2=ESInvcolor,shape=21) %>% rbind(forest2nat)

ecoreglabels <- c("Boreal Forests/Taiga", "Deserts and Xeric\nShrublands", "Flooded Grasslands\nand Savannas", "Mediterranean Forests,\nWoodlands and Scrub", "Montane Grasslands\nand Shrublands","Temperate Broadleaf\nand Mixed Forests","Temperate Conifer\nForests","Temperate Grasslands,\nSavannas and Shrublands", "Tropical and Subtropical\nDry Broadleaf Forests", "Tropical and Subtropical\nGrasslands, Savannas\nand Shrublands", "Tropical and Subtropical\nMoist Broadleaf Forests","","","","","","")

p2<-
  ggplot(data=forest2, 
       aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`)) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  geom_errorbarh(data=forest2, height=.1, size=1, color=forest2$treatment2)+
  geom_point(data=forest2, size=4, shape=forest2$shape, color=forest2$treatment2, aes(fill=treatment), stroke=1)+
  scale_y_continuous(name = "", breaks=1:nrow(forest2),trans="reverse",labels = ecoreglabels) +
  scale_fill_manual(values = c(ESNatcolor, 'white', ESInvcolor,'white'))+
  xlab("Effect size (ES, mean + 95%PI)") + # I might not need this when stacking them all
  coord_cartesian(ylim = c(11,1), xlim=c(-5.5,4.7), clip = "off")+
  theme_bw()+ 
  theme(text=element_text(size=13, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.text.y = element_text(size=9),
      axis.title.x = element_blank(),
      plot.margin = margin(1,10,1,1),
      axis.line=element_line(color="black"),
      panel.border = element_blank()) + 
  geom_text_repel(aes(label = n), size=3, color=forest2$treatment2) +
  annotate("text", x=-5.5, y=0.63, label="(b)", color="black", fontface = "bold", size=4)+
  ggh4x::force_panelsizes(cols = unit(3, "in"), rows = unit(4, "in"))
```

## Forest community
### Invasive performance
#### organizing model data
```{r}
InvAllDataModel2 <- InvAllDataModel %>% filter(!is.na(ForestCommunity),
                                               ForestCommunity!="not_reported")%>% 
  mutate(ForestCommunity=ifelse(ForestCommunity=="riparian_broadleaf" | ForestCommunity=="riparian_deciduous", "riparian", ForestCommunity),
         ForestCommunity=ifelse(ForestCommunity=="conifer_mixed", "conifer", ForestCommunity)) %>% 
  group_by(ForestCommunity) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(InvAllDataModel2)
StudyID <- InvAllDataModel2$StudyID
CategoryID <- InvAllDataModel2$CategoryID
invESavemean <- InvAllDataModel2$ESInvmean
invESavesd <- InvAllDataModel2$ESInvSD
M <- max(InvAllDataModel2$ESInvSD, na.rm = T) 
K <- length(unique(InvAllDataModel2$CategoryID))
R <- K+1 
W <- length(unique(InvAllDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, invESavemean=invESavemean, invESavesd=invESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(invESavesd[i]*invESavesd[i])

invESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
invESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ # SRE
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials 
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "invESavemean.h", "tau", "var")

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 9, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESInv-ForestComm.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) 
invESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^invESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(invESavemean.h,invESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESInvmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 1.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- InvAllDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, ForestCommunity, n) %>% distinct() 

         
inv <- as.data.frame(es.simCat[["summary"]])
inv <- inv %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "invasive")

forest3inv <- inv %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
```

### Native performance
#### organizing model data
```{r}
BioticDataModel2 <- BioticDataModel %>% filter(!is.na(ForestCommunity),
                                               ForestCommunity!="not_reported")%>% 
  mutate(ForestCommunity=ifelse(ForestCommunity=="riparian_broadleaf" | ForestCommunity=="riparian_deciduous", "riparian", ForestCommunity)) %>% 
  group_by(ForestCommunity) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(BioticDataModel2)
StudyID <- BioticDataModel2$StudyID
CategoryID <- BioticDataModel2$CategoryID
natESavemean <- BioticDataModel2$ESNatmean
natESavesd <- BioticDataModel2$ESNatSD
M <- max(BioticDataModel2$ESNatSD, na.rm = T) 
K <- length(unique(BioticDataModel2$CategoryID))
R <- K+1 
W <- length(unique(BioticDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, natESavemean=natESavemean, natESavesd=natESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
natESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(natESavesd[i]*natESavesd[i])

natESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
natESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ # SRE
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "natESavemean.h", "tau", "var") 

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 15, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESNat-ForestComm.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]])
natESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^natESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(natESavemean.h,natESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESNatmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 0.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- BioticDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, ForestCommunity, n) %>% distinct() 

         
nat <- as.data.frame(es.simCat[["summary"]])
nat <- nat %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "native")

forest3nat <- nat %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
  
```

```{r echo=FALSE}
# forest plot of forest community, with invasive species and native community ES
forest3nat <- forest3nat %>% mutate(index=ifelse(index<4, index+0.2,
                                       ifelse(index>3&index<12, index+1.2, 
                                              ifelse(ForestCommunity=="woodland", 14.2, index))),
                                 treatment=ifelse(CategoryID==5, "a", "b"),
                                 treatment2=ESNatcolor,shape=23)

forest3 <- forest3inv %>% mutate(treatment=ifelse(CategoryID==2|
                                                    CategoryID==4|
                                                    CategoryID==9|
                                                    CategoryID==13, "d", "c"), 
                                   treatment2=ESInvcolor,shape=21) %>% rbind(forest3nat)

forestlabels <-  c("Broadleaf", "Conifer", "Deciduous", "Deciduous floodplain", "Dry", "Evergreen", "Floodplain","Mixed","Riparian","Savanna","Semi-deciduous","Semi-dry", "Shrubland","Woodland","","","","","","","","","","","")

p3<-
  ggplot(data=forest3, 
       aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`)) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  geom_errorbarh(data=forest3, height=.1, size=1, color=forest3$treatment2)+
  geom_point(data=forest3, size=4, shape=forest3$shape, color=forest3$treatment2, aes(fill=treatment), stroke=1)+
  scale_y_continuous(name = "", breaks=1:nrow(forest3),trans="reverse",labels = forestlabels) +
  scale_fill_manual(values = c(ESNatcolor, 'white', ESInvcolor,'white'))+
  xlab("Effect size (ES, mean + 95%PI)") + # I might not need this when stacking them all
  coord_cartesian(ylim = c(14.3,0.9), xlim=c(-6.5,5), clip = "off")+
  theme_bw()+
  theme(text=element_text(size=13, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title.x = element_blank(),
      plot.margin = margin(1,1,1,1),
      axis.line=element_line(color="black"),
      panel.border = element_blank()) + 
  geom_text_repel(aes(label = n), size=3, color=forest3$treatment2) +
  annotate("text", x=-6.5, y=0.45, label="(c)", color="black", fontface = "bold", size=4)+
  ggh4x::force_panelsizes(rows = unit(4, "in"), cols = unit(3, "in"))
```

## Forest type
### Invasive performance
#### organizing model data
```{r}
InvAllDataModel2 <- InvAllDataModel %>% filter(!is.na(TypeOfForest),
                                               TypeOfForest!="not_reported") %>% 
  group_by(TypeOfForest) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(InvAllDataModel2)
StudyID <- InvAllDataModel2$StudyID
CategoryID <- InvAllDataModel2$CategoryID
invESavemean <- InvAllDataModel2$ESInvmean
invESavesd <- InvAllDataModel2$ESInvSD
M <- max(InvAllDataModel2$ESInvSD, na.rm = T) 
K <- length(unique(InvAllDataModel2$CategoryID))
R <- K+1 
W <- length(unique(InvAllDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, invESavemean=invESavemean, invESavesd=invESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(invESavesd[i]*invESavesd[i])

invESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
invESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ 
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "invESavemean.h", "tau","var")

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value: 
```{r}
es.simCat[["DIC"]]
```

+ Trace plots

```{r}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 10, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESInv-ForestType.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) 
invESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^invESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(invESavemean.h,invESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESInvmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 1.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- InvAllDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, TypeOfForest, n) %>% distinct() 

         
inv <- as.data.frame(es.simCat[["summary"]])
inv <- inv %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "invasive")

forest4inv <- inv %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
```

### Native performance
#### organizing model data
```{r}
BioticDataModel2 <- BioticDataModel %>% filter(!is.na(TypeOfForest),
                                               TypeOfForest!="not_reported") %>% 
  group_by(TypeOfForest) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(BioticDataModel2)
StudyID <- BioticDataModel2$StudyID
CategoryID <- BioticDataModel2$CategoryID
natESavemean <- BioticDataModel2$ESNatmean
natESavesd <- BioticDataModel2$ESNatSD
M <- max(BioticDataModel2$ESNatSD, na.rm = T) 
K <- length(unique(BioticDataModel2$CategoryID))
R <- K+1 
W <- length(unique(BioticDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, natESavemean=natESavemean, natESavesd=natESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
natESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(natESavesd[i]*natESavesd[i])

natESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
natESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ 
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "natESavemean.h", "tau","var") 

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 16, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESNat-ForestType.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) 
natESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^natESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(natESavemean.h,natESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESNatmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 0.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- BioticDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, TypeOfForest, n) %>% distinct() 

         
nat <- as.data.frame(es.simCat[["summary"]])
nat <- nat %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "native")

forest4nat <- nat %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
```

```{r echo=FALSE}
forest4nat <- forest4nat %>% mutate(index=ifelse(TypeOfForest=="natural", 1.2,
                                       ifelse(TypeOfForest=="natural_secondary", 3.2,
                                       ifelse(TypeOfForest=="natural_secondary_invaded_canopy", 4.2,
                                       ifelse(TypeOfForest=="plantation", 5.2,
                                       ifelse(TypeOfForest=="restoration_planted", 7.2, index))))),
                                 treatment=ifelse(CategoryID==4, "a", "b"),
                                 treatment2=ESNatcolor,shape=23)

forest4 <- forest4inv %>% 
  mutate(treatment=ifelse(CategoryID>3, "d", "c"), 
                                   treatment2=ESInvcolor,shape=21) %>% 
  rbind(forest4nat)

forestlabels <-  c("Natural", "Natural primary", "Natural secondary", "Natural secondary\nwith invaded canopy", "Plantation", "Plantation and natural", "Restoration (planted)","","","","","")


p4<-
  ggplot(data=forest4, 
       aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`)) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  geom_errorbarh(data=forest4, height=.1, size=1, color=forest4$treatment2)+
  geom_point(data=forest4, size=4, shape=forest4$shape, color=forest4$treatment2, aes(fill=treatment), stroke=1)+
  scale_y_continuous(name = "", breaks=1:nrow(forest4),trans="reverse",labels = forestlabels) +
  scale_fill_manual(values = c(ESNatcolor, 'white', ESInvcolor,'white'))+
  xlab("Effect size (ES, mean + 95%PI)") + 
  coord_cartesian(ylim = c(7.3,0.9), xlim=c(-6.3,6.5), clip = "off")+
  theme_bw()+ 
  theme(text=element_text(size=13, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title.x = element_blank(),
      plot.margin = margin(1,1,1,1),
      axis.line=element_line(color="black"),
      panel.border = element_blank()) +
  geom_text_repel(aes(label = n), size=3, color=forest4$treatment2) +
  annotate("text", x=-6.4, y=0.7, label="(d)", color="black", fontface = "bold", size=4)+
  ggh4x::force_panelsizes(cols = unit(3, "in"), rows = unit(4, "in"))
```

## Study type
### Invasive performance
#### organizing model data
```{r}
InvAllDataModel2 <- InvAllDataModel %>% 
  group_by(TypeOfStudy) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(InvAllDataModel2)
StudyID <- InvAllDataModel2$StudyID
CategoryID <- InvAllDataModel2$CategoryID
invESavemean <- InvAllDataModel2$ESInvmean
invESavesd <- InvAllDataModel2$ESInvSD
M <- max(InvAllDataModel2$ESInvSD, na.rm = T) 
K <- length(unique(InvAllDataModel2$CategoryID))
R <- K+1 
W <- length(unique(InvAllDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, invESavemean=invESavemean, invESavesd=invESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(invESavesd[i]*invESavesd[i])

invESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
invESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ 
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials 
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "invESavemean.h", "tau","var") 

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 11, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESInv-StudyType.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]])
invESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^invESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(invESavemean.h,invESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESInvmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 1.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- InvAllDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, TypeOfStudy, n) %>% distinct() 

         
inv <- as.data.frame(es.simCat[["summary"]])
inv <- inv %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "invasive")

forest1inv <- inv %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
```


### Native performance
#### organizing model data
```{r}
BioticDataModel2 <- BioticDataModel %>% 
  group_by(TypeOfStudy) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(BioticDataModel2)
StudyID <- BioticDataModel2$StudyID
CategoryID <- BioticDataModel2$CategoryID
natESavemean <- BioticDataModel2$ESNatmean
natESavesd <- BioticDataModel2$ESNatSD
M <- max(BioticDataModel2$ESNatSD, na.rm = T) 
K <- length(unique(BioticDataModel2$CategoryID))
R <- K+1 
W <- length(unique(BioticDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, natESavemean=natESavemean, natESavesd=natESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
natESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(natESavesd[i]*natESavesd[i])

natESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
natESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ 
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "natESavemean.h", "tau", "var") 

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 17, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESNat-StudyType.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) 
natESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^natESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(natESavemean.h,natESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESNatmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 0.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- BioticDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, TypeOfStudy, n) %>% distinct() 

         
nat <- as.data.frame(es.simCat[["summary"]])
nat <- nat %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "native")

forest1nat <- nat %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
```

```{r echo=FALSE}
# forest plot of study type, with invasive species and native community ES
forest1inv <- forest1inv %>% mutate(treatment="c", treatment2=ESInvcolor, shape=21)
forest1 <- forest1nat %>% mutate(index=index+0.2,
                                 treatment=ifelse(CategoryID==3, "b", "a"),
                                 treatment2=ESNatcolor,shape=23) %>% rbind(forest1inv)
p1<-
ggplot(data=forest1, 
       aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`)) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  geom_errorbarh(data=forest1, height=.1, size=1, color=forest1$treatment2)+
  geom_point(data=forest1, size=4, shape=forest1$shape, color=forest1$treatment2, aes(fill=treatment), stroke=1)+
  scale_y_continuous(name = "", breaks=1:nrow(forest1),trans="reverse",
                      labels = c("Field experiment", "Greenhouse", "Observational", "",  "", "")
                     ) +
  scale_fill_manual(values = c('white', ESNatcolor, ESInvcolor))+
  coord_cartesian(ylim = c(3.3,0.9), xlim=c(-1.7,2.5), clip = "off")+
  theme_bw()+ 
  theme(text=element_text(size=13, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title.x = element_blank(), # I might need this when stacking them all
      plot.margin = margin(1,1,1,1),
      axis.line=element_line(color="black"),
      panel.border = element_blank()) +  
  geom_text_repel(aes(label = n), size=3, color=forest1$treatment2)  +
  annotate("text", x=-1.73, y=0.82, label="(a)", color="black", fontface = "bold", size=4)+
  ggh4x::force_panelsizes(cols = unit(3, "in"), rows=unit(4, "in"))
```

## Disturbance
### Invasive performance
#### organizing model data
```{r}
InvAllDataModel2 <- InvAllDataModel %>% filter(!is.na(Disturbance), Disturbance!="not_reported") %>% 
  group_by(Disturbance) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(InvAllDataModel2)
StudyID <- InvAllDataModel2$StudyID
CategoryID <- InvAllDataModel2$CategoryID
invESavemean <- InvAllDataModel2$ESInvmean
invESavesd <- InvAllDataModel2$ESInvSD
M <- max(InvAllDataModel2$ESInvSD, na.rm = T) 
K <- length(unique(InvAllDataModel2$CategoryID))
R <- K+1 
W <- length(unique(InvAllDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, invESavemean=invESavemean, invESavesd=invESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(invESavesd[i]*invESavesd[i])

invESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
invESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ 
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "invESavemean.h", "tau","var") 

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 12, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESInv-Dist.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) 
invESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^invESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(invESavemean.h,invESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESInvmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 1.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- InvAllDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, Disturbance, n) %>% distinct() 

         
inv <- as.data.frame(es.simCat[["summary"]])
inv <- inv %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "invasive")

forest5inv <- inv %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
```

### Native performance
#### organizing the data
```{r}
BioticDataModel2 <- BioticDataModel %>% filter(!is.na(Disturbance), Disturbance!="not_reported") %>% 
  group_by(Disturbance) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(BioticDataModel2)
StudyID <- BioticDataModel2$StudyID
CategoryID <- BioticDataModel2$CategoryID
natESavemean <- BioticDataModel2$ESNatmean
natESavesd <- BioticDataModel2$ESNatSD
M <- max(BioticDataModel2$ESNatSD, na.rm = T) 
K <- length(unique(BioticDataModel2$CategoryID))
R <- K+1 
W <- length(unique(BioticDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, natESavemean=natESavemean, natESavesd=natESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
natESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(natESavesd[i]*natESavesd[i])

natESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
natESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ 
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials 
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "natESavemean.h", "tau", "var") 

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 18, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESNat-Dist.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]])
natESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^natESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(natESavemean.h,natESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESNatmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 0.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- BioticDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, Disturbance, n) %>% distinct() 

         
nat <- as.data.frame(es.simCat[["summary"]])
nat <- nat %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "native")

forest5nat <- nat %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
```

```{r echo=FALSE}
# forest plot of disturbance, with invasive species and native community ES
forest5inv <- forest5inv %>% mutate(treatment="c", treatment2=ESInvcolor, shape=21)
forest5 <- forest5nat %>% mutate(index=index+0.2,
                                 treatment=ifelse(CategoryID==2, "b", "a"),
                                 treatment2=ESNatcolor, shape=23) %>% rbind(forest5inv)

p5<-
  ggplot(data=forest5, 
       aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`)) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  geom_errorbarh(data=forest5, height=.1, size=1, color=forest5$treatment2)+  
  geom_point(data=forest5, size=4, shape=forest5$shape, color=forest5$treatment2, aes(fill=treatment), stroke=1)+
  scale_y_continuous(name = "", breaks=1:nrow(forest5),trans="reverse",
                     labels = c("Disturbance\nabsent","Disturbance\npresent","","")) +
  scale_fill_manual(values = c('white', ESNatcolor, ESInvcolor))+
  xlab("Effect size (ES, mean + 95%PI)")+
   coord_cartesian(ylim = c(2.3,0.9), xlim=c(-0.55,1.05), clip = "off")+
  theme_bw()+ 
  theme(text=element_text(size=13, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title.x = element_text(size = 10),
      plot.margin = margin(1,1,1,1),
      axis.line=element_line(color="black"),
      panel.border = element_blank())+
  geom_text_repel(aes(label = n), size=3, color=forest5$treatment2) +
  annotate("text", x=-0.55, y=0.85, label="(e)", color="black", fontface = "bold", size=4, alpha=1)+
  ggh4x::force_panelsizes( cols = unit(3, "in"), rows = unit(4, "in"))
```

## Type of disturbance
### Invasive performance
#### organizing model data
```{r}
InvAllDataModel2 <- InvAllDataModel %>% filter(!is.na(NatureOfDisturbance), NatureOfDisturbance!="not_reported") %>% 
  group_by(NatureOfDisturbance) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(InvAllDataModel2)
StudyID <- InvAllDataModel2$StudyID
CategoryID <- InvAllDataModel2$CategoryID
invESavemean <- InvAllDataModel2$ESInvmean
invESavesd <- InvAllDataModel2$ESInvSD
M <- max(InvAllDataModel2$ESInvSD, na.rm = T) 
K <- length(unique(InvAllDataModel2$CategoryID))
R <- K+1 
W <- length(unique(InvAllDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, invESavemean=invESavemean, invESavesd=invESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(invESavesd[i]*invESavesd[i])

invESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
invESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ 
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "invESavemean.h", "tau","var") 

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 13, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESInv-DistType.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]])
invESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^invESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(invESavemean.h,invESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESInvmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 1.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- InvAllDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, NatureOfDisturbance, n) %>% distinct() 

         
inv <- as.data.frame(es.simCat[["summary"]])
inv <- inv %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "invasive")

forest6inv <- inv %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
```

### Native performance
#### organizing model data
```{r}
BioticDataModel2 <- BioticDataModel %>% filter(!is.na(NatureOfDisturbance), NatureOfDisturbance!="not_reported") %>% 
  group_by(NatureOfDisturbance) %>% mutate(CategoryID = cur_group_id()) %>% ungroup() %>% 
  group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup()

J <- nrow(BioticDataModel2)
StudyID <- BioticDataModel2$StudyID
CategoryID <- BioticDataModel2$CategoryID
natESavemean <- BioticDataModel2$ESNatmean
natESavesd <- BioticDataModel2$ESNatSD
M <- max(BioticDataModel2$ESNatSD, na.rm = T) 
K <- length(unique(BioticDataModel2$CategoryID))
R <- K+1 
W <- length(unique(BioticDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, CategoryID=CategoryID, natESavemean=natESavemean, natESavesd=natESavesd, M=M, K=K, R=R, W=W)
```

#### model
```{r}
modelESCat <- function() {
  for(i in 1:J){
# missing values
natESavesd[i]~dnorm(M,1)%_%C(0,)

tauI[i]<-1/(natESavesd[i]*natESavesd[i])

natESavemean[i]~dnorm(iE[i],tau[CategoryID[i]])
natESavemean.h[i]~dnorm(iE[i],tau[CategoryID[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[CategoryID[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
  for(i in 1:K){
beta[i]~dnorm(0,0.0001)
  }

for(i in 1:W){ 
iSRE[i]~dnorm(0,tau[R])
}
  
  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,10)
}

} # end model

# initials
inits <- list(
  list(beta = c(rep(1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(-1, times=K)), var = c(rep(1, times = R))),
  list(beta = c(rep(0, times=K)), var = c(rep(1, times = R)))
)

# parameters
parameters <- c("beta", "natESavemean.h", "tau","var") 

# running model
es.simCat <- bugs(datamodel, inits, parameters, model.file = modelESCat, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simCat[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simCat, "beta")
```

+ Model parameters
```{r parameters 19, echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname))
write.csv(x, file.path(parametersout, "ESNat-DistType.csv"), row.names = F)
x
```

+ Model fit

```{r echo=FALSE}
x<-as.data.frame(es.simCat[["summary"]]) 
natESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^natESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(natESavemean.h,natESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESNatmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 0.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation to plot data
samplesize <- BioticDataModel2 %>% 
  group_by(CategoryID) %>%
  add_tally() %>% ungroup() %>% select(CategoryID, NatureOfDisturbance, n) %>% distinct() 

         
nat <- as.data.frame(es.simCat[["summary"]])
nat <- nat %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(grepl("^beta", rowname)) %>%
  mutate(type = "native")

forest6nat <- nat %>%
  arrange(rowname) %>%
  mutate(comparison = as.integer(str_sub(rowname, 6, 6)),
         comparison1 = as.integer(str_sub(rowname,6, 7)),
         CategoryID = ifelse(is.na(comparison1), comparison, comparison1)) %>% 
  mutate_all(na_if,"") %>% 
  arrange(CategoryID) %>%
  left_join(samplesize) %>%
  mutate(index = ifelse(is.na(CategoryID), nrow(inv), CategoryID),
         n = paste0('(', ifelse(is.na(n), sum(n, na.rm=TRUE), n), ')')) 
```

```{r echo=FALSE}
# forest plot of type of disturbance, with invasive species and native community ES
forest6inv <- forest6inv %>% mutate(treatment=ifelse(CategoryID==3, "d", "c"), treatment2=ESInvcolor, shape=21)
forest6 <- forest6nat %>% mutate(index=index+0.2,
                                 treatment=ifelse(CategoryID==2, "a", "b"),
                                 treatment2=ESNatcolor, shape=23) %>% rbind(forest6inv)

p6<-
  ggplot(data=forest6,
       aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`)) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  geom_errorbarh(data=forest6, height=.1, size=1, color=forest6$treatment2)+
  geom_point(data=forest6, size=4, shape=forest6$shape, color=forest6$treatment2, aes(fill=treatment), stroke=1)+
  scale_y_continuous(name = "", breaks=1:nrow(forest6),trans="reverse", 
                     labels = c("Natural\ndisturbance","Human-induced\ndisturbance","Both types of\ndisturbance","","",""))+
  scale_fill_manual(values = c(ESNatcolor, 'white', ESInvcolor,'white')) +
  xlab("Effect size (ES, mean + 95%PI)") + 
  coord_cartesian(ylim = c(3.3,0.9), xlim=c(-2.2,1.5), clip = "off")+
  theme_bw()+ 
  theme(text=element_text(size=13, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title.x = element_text(size = 10),
      plot.margin = margin(1,1,1,1),
      axis.line=element_line(color="black"),
      panel.border = element_blank()) + 
  geom_text_repel(aes(label = n), size=3, color=forest6$treatment2) +
  annotate("text", x=-2.25, y=0.82, label="(f)", color="black", fontface = "bold", size=4)+
  ggh4x::force_panelsizes(cols = unit(3, "in"), rows = unit(4, "in"))
```


```{r include=FALSE}
# overall figure
p1 <- ggplotGrob(p1)
p2 <- ggplotGrob(p2)
p3 <- ggplotGrob(p3)
p4 <- ggplotGrob(p4)
p5 <- ggplotGrob(p5)
p6 <- ggplotGrob(p6)

#showing them all
png(file.path(graph.wd, "ESNatESInv_variables.png"),width=9.4, height=13, res=300,units="in")
grid::grid.draw(cbind(rbind(p1,p3,p5), rbind(p2,p4,p6)))
dev.off()
```

# **SM: how does invasive species and native community performance vary with disturbance across mechanisms?**
## Invasive species
### data preparation
```{r echo=TRUE}
# adding information from file with the estimated ES values (with sd) and calculating ES for the observations with no associated variability
InvAllDataModeltemp <- InvAllDataModel %>% mutate(Disturbance=ifelse(Disturbance=="yes", 1,2), # categories "no disturbance" and "not reported" are aggregated here
                                               Disturbance=ifelse(is.na(Disturbance),2,Disturbance),
                                              Type=ifelse(NatureOfDisturbance=="not_reported"|is.na(NatureOfDisturbance),4,NatureOfDisturbance),
                                              Type=as.integer(Type),
                                              Type=ifelse(Disturbance==2,5,Type))  

InvAllDataModel2 <- InvAllDataModeltemp %>% 
  group_by(Mechanism, Disturbance) %>%
  add_tally() %>% 
  group_by(Mechanism, Disturbance, Type) %>%
  add_tally() %>% filter(n>1,nn>1) %>% group_by(StudyID) %>% mutate(StudyID = cur_group_id()) %>% ungroup() %>% select(-n, -nn)

# Observations with one sample size
InvAllDataModelEXCLUDED <- InvAllDataModeltemp %>% 
  group_by(Mechanism, Disturbance) %>%
  add_tally() %>% 
  group_by(Mechanism, Disturbance, Type) %>%
  add_tally() %>% filter(n==1|nn==1) %>% 
  select(Mechanism, Disturbance, Type, ESInvmean, ESInvSD) %>% filter(!is.na(ESInvSD))
```

#### organizing model data
```{r}
J <- nrow(InvAllDataModel2)
StudyID <- InvAllDataModel2$StudyID 
Mechanism <- InvAllDataModel2$Mechanism
Disturbance <- InvAllDataModel2$Disturbance
Type <- InvAllDataModel2$Type 
invESavemean <- InvAllDataModel2$ESInvmean
invESavesd <- InvAllDataModel2$ESInvSD
M <- max(InvAllDataModel2$ESInvSD, na.rm = T) 
K <- length(unique(InvAllDataModel2$Mechanism))
R <- 1 
W <- length(unique(InvAllDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, Mechanism=Mechanism, Disturbance=Disturbance, 
                   Type=Type, invESavemean=invESavemean, invESavesd=invESavesd, 
                   M=M, R=R, W=W, K=K)
```

#### model 
```{r}
# hierarchical model
modelESInv <- function() {
  for(i in 1:J){
# missing values
invESavesd[i]~dnorm(M,1)%_%C(0,) 

tauI[i]<-1/(invESavesd[i]*invESavesd[i])

invESavemean[i]~dnorm(iE[i],tauAll[Mechanism[i]])
invESavemean.h[i]~dnorm(iE[i],tauAll[Mechanism[i]])
ie[i]~dnorm(0,tauI[i])
iE[i]<-beta[Mechanism[i],Disturbance[i],Type[i]]+iSRE[StudyID[i]]+ie[i]

}

# priors
    beta[1,1,1]~dnorm(BB[1,1],tauBB[1,1])
    beta[1,1,2]~dnorm(BB[1,1],tauBB[1,1])
    beta[1,1,3]<-0 #no data (sample size 1)
    beta[1,1,4]<-0 #no data
    beta[1,2,1]<-0 #no data
    beta[1,2,2]<-0 #no data
    beta[1,2,3]<-0 #no data
    beta[1,2,4]<-0 #no data
    beta[1,2,5]<-0 #no data (sample size 1)
    beta[2,1,1]~dnorm(BB[2,1],tauBB[2,1])
    beta[2,1,2]~dnorm(BB[2,1],tauBB[2,1])
    beta[2,1,3]~dnorm(BB[2,1],tauBB[2,1])
    beta[2,1,4]~dnorm(BB[2,1],tauBB[2,1])
    beta[2,2,1]<-0 #no data
    beta[2,2,2]<-0 #no data
    beta[2,2,3]<-0 #no data
    beta[2,2,4]<-0 #no data
    beta[2,2,5]<-BB[2,2] #single subcategory
    beta[3,1,1]~dnorm(BB[3,1],tauBB[3,1])
    beta[3,1,2]~dnorm(BB[3,1],tauBB[3,1])
    beta[3,1,3]~dnorm(BB[3,1],tauBB[3,1])
    beta[3,1,4]~dnorm(BB[3,1],tauBB[3,1])
    beta[3,2,1]<-0 #no data
    beta[3,2,2]<-0 #no data
    beta[3,2,3]<-0 #no data
    beta[3,2,4]<-0 #no data
    beta[3,2,5]<-BB[3,2] #single subcategory
    
    BB[1,1]~dnorm(B[1],tauB[1])
    tauBB[1,1]<-1/varBB[1,1]
    varBB[1,1]~dunif(0,1)
    
    BB[1,2]<-0
    
    BB[2,1]~dnorm(B[2],tauB[2])
    tauBB[2,1]<-1/varBB[2,1]
    varBB[2,1]~dunif(0,1)

    BB[2,2]~dnorm(B[2],tauB[2])

    BB[3,1]~dnorm(B[3],tauB[3])
    tauBB[3,1]<-1/varBB[3,1]
    varBB[3,1]~dunif(0,1)

    BB[3,2]~dnorm(B[3],tauB[3])

B[1]~dnorm(0,0.1)
tauB[1]<-1/varB[1]
varB[1]~dunif(0,1)

B[2]~dnorm(0,0.1)
tauB[2]<-1/varB[2]
varB[2]~dunif(0,1)

B[3]~dnorm(0,0.1)
tauB[3]<-1/varB[3]
varB[3]~dunif(0,1)

for(i in 1:W){ 
iSRE[i]~dnorm(0,tau[R])
}

  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,1)
}

    for(i in 1:K){
tauAll[i]<-1/varAll[i]
varAll[i]~dunif(0,1)
    }

} # end model

# initials 
inits <- list(
  list(beta = c(rep(2, times=12)),
       var = c(rep(0.5, times=R)), 
       varBB=c(rep(0.5,times=3)),
       varB=c(rep(0.5,times=K)), 
       varAll=c(rep(1,times=K))),
  
  list(beta = c(rep(-1, times=12)), 
       var = c(rep(0.5, times=R)),
       varBB=c(rep(0.5,times=3)), 
       varB=c(rep(0.5,times=K)), 
       varAll=c(rep(1,times=K))),
  
  list(beta = c(rep(0, times=12)), 
       var = c(rep(0.5, times=R)),
       varBB=c(rep(0.5,times=3)), 
       varB=c(rep(0.5,times=K)), 
       varAll=c(rep(1,times=K)))
)

# parameters
parameters <- c("B", "BB", "beta", "invESavemean.h", "tau", 
                "tauAll", "tauB", "tauBB", "varB", "varBB", "var")
# running model
es.simInvDist <- bugs(datamodel, inits, parameters, model.file = modelESInv, n.chains=3, n.burnin=25000, n.iter=75000, debug = F)
```

+ DIC value:
```{r echo=FALSE}
es.simInvDist[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simInvDist, "beta")
traplot(es.simInvDist, "B")
traplot(es.simInvDist, "BB")
```

+ Model parameters
```{r parameters 6, echo=FALSE}
x<-as.data.frame(es.simInvDist[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname)|
                                                     grepl("^B", rowname))
write.csv(x, file.path(parametersout, "ESInv-Disturbance.csv"), row.names = F)
x
```

+ Model fit:

```{r echo=FALSE}
x<-as.data.frame(es.simInvDist[["summary"]]) 
invESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^invESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(invESavemean.h,invESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESInvmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = 0.2, y = 1.5, labels = mylabel)

```

```{r include=FALSE}
# data preparation for plot (invasive ES)
samplesize2 <- InvAllDataModel2 %>% 
  group_by(Mechanism) %>%
  add_tally() %>% ungroup() %>% mutate(Disturbance=NA, Type=NA) %>% 
  select(Mechanism, Disturbance, Type, n) %>% distinct() %>% arrange(Mechanism) %>% 
  mutate(index=c(1,6,13),
         size=14,
         label=c("Propagule pressure", "Empty niches", "Low biotic resistance"),
         n=ifelse(index==1, n+1, n),shape=21,sizedot=6,treatment=c("a","a","a"))  

samplesize3 <- InvAllDataModel2 %>% 
  group_by(Mechanism, Disturbance) %>%
  add_tally() %>% ungroup() %>% mutate(Type=NA) %>% 
  select(Mechanism, Disturbance, Type, n) %>% distinct() %>% arrange(Mechanism, Disturbance) %>%
  mutate(index=c(3,8,7,15,14),shape=21,sizedot=5,size=12,treatment=c("a","b","b","b","b"),
         label=c("Disturbance","Disturbance","No disturbance","Disturbance","No disturbance"))

samplesize <- InvAllDataModel2 %>% 
  group_by(Mechanism, Disturbance, Type) %>%
  add_tally() %>% ungroup() %>% select(Mechanism, Disturbance, Type, n) %>% distinct() %>% 
 arrange(Mechanism, Disturbance, Type) %>% 
  mutate(index=c(4,5,9,10,11,12,NA,16,17,18,19, NA),size=10,
         label=c("Natural","Human","Natural","Human","Both","Not reported",NA,"Natural","Human","Both","Not reported",NA),shape=21,sizedot=4,treatment=c("b","b","b","b","b","b",NA,"b","b","b","b",NA)) %>% 
  filter(!is.na(index)) %>% 
 rbind(samplesize2) %>% rbind(samplesize3) %>% mutate(n = paste0('(', n, ')')) 

ESInvEXCLUDED <- InvAllDataModelEXCLUDED %>% mutate(rowname="BB[1,2]",name="BB", `2.5%`=ESInvmean-(ESInvSD*2),
                                                    `97.5%`=ESInvmean+(ESInvSD*2),
                                                    n = paste0('(', 1, ')'),
                                                    index=2,size=12,shape=21,sizedot=5,treatment="a",
                                                    label="No disturbance") %>% 
  rename(mean=ESInvmean, sd=ESInvSD)
         
inv <- as.data.frame(es.simInvDist[["summary"]])
forestinv <- inv %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(!grepl("^invESavemean.h", rowname),
         !grepl("^tau", rowname),
         !grepl("^var", rowname),
         rowname != "deviance") %>% 
  separate(col=rowname, remove=F, into=c("name","Mechanism","Disturbance","Type")) %>% 
  mutate(Disturbance=as.numeric(Disturbance),
         Mechanism=as.numeric(Mechanism),
         Type=as.numeric(Type)) %>% 
  mutate_all(na_if,"") %>% 
  left_join(samplesize) %>% 
  filter(!is.na(index)) %>% 
  rbind(ESInvEXCLUDED) %>% 
  arrange(index) %>% 
  mutate(treatment2=ESInvcolor)
```


## Native community
### data preparation
```{r}
# adding information from file with the estimated ES values (with sd) and calculating ES for the observations with no associated variability
NatAllDataModeltemp <- data %>% filter(ObsIDOriginal %in% NatUniqueBiotic$ObsIDOriginal) %>% 
  left_join(NatBioticEstimate, by = "ObsIDOriginal") %>%
  mutate(ESNatmean = ifelse(is.na(ESNatmean), ((NatTreatmean-NatContmean)/((NatTreatmean+NatContmean)/2)), ESNatmean),
         ObsID = row_number()) %>% mutate(Disturbance=ifelse(Disturbance=="yes", 1,2), # categories "no disturbance" and "not reported" are aggregated here
                                               Disturbance=ifelse(is.na(Disturbance),2,Disturbance),
                                              Type=ifelse(NatureOfDisturbance=="not_reported"|is.na(NatureOfDisturbance),4,NatureOfDisturbance),
                                              Type=as.integer(Type),
                                              Type=ifelse(Disturbance==2,5,Type)) 

NatAllDataModel2 <- NatAllDataModeltemp %>% 
  group_by(Mechanism, Disturbance) %>%
  add_tally() %>% 
  group_by(Mechanism, Disturbance, Type) %>%
  add_tally() %>% filter(n>1,nn>1) %>% group_by(StudyIDOriginal) %>% mutate(StudyID = cur_group_id()) %>% ungroup() %>% select(-n, -nn)

NatAllDataModelEXCLUDED <- NatAllDataModeltemp %>% 
  group_by(Mechanism, Disturbance) %>%
  add_tally() %>% 
  group_by(Mechanism, Disturbance, Type) %>%
  add_tally() %>% filter(n==1|nn==1) %>% # two observations with a single sample size. will include the one with reported sd later in the graph
  select(Mechanism, Disturbance, Type, ESNatmean, ESNatSD) %>% filter(!is.na(ESNatSD)) %>% ungroup() 
```

#### organizing model data
```{r}
J <- nrow(NatAllDataModel2)
StudyID <- NatAllDataModel2$StudyID 
Mechanism <- NatAllDataModel2$Mechanism
Disturbance <- NatAllDataModel2$Disturbance
Type <- NatAllDataModel2$Type # disturbance type
natESavemean <- NatAllDataModel2$ESNatmean
natESavesd <- NatAllDataModel2$ESNatSD
M <- max(NatAllDataModel2$ESNatSD, na.rm = T) 
K <- length(unique(NatAllDataModel2$Mechanism))
R <- 1  
W <- length(unique(NatAllDataModel2$StudyID))

datamodel <- list (J=J, StudyID=StudyID, Mechanism=Mechanism, Disturbance=Disturbance, Type=Type, natESavemean=natESavemean, natESavesd=natESavesd, M=M, R=R, W=W, K=K)
```

#### model 
```{r}
# hierarchical model
modelESNat <- function() {
  for(i in 1:J){
# missing values
natESavesd[i]~dnorm(M,1)%_%C(0,) # M=maximum variance

tauN[i]<-1/(natESavesd[i]*natESavesd[i])

natESavemean[i]~dnorm(nE[i],tauAll[Mechanism[i]])
natESavemean.h[i]~dnorm(nE[i],tauAll[Mechanism[i]])
ne[i]~dnorm(0,tauN[i])
nE[i]<-beta[Mechanism[i],Disturbance[i],Type[i]]+iSRE[StudyID[i]]+ne[i]

}

# priors
    beta[1,1,1]<-0 #no data
    beta[1,1,2]<-B[1]
    beta[1,1,3]<-0 #no data (sample size 1)
    beta[1,1,4]<-0 #no data
    beta[1,2,1]<-0 #no data
    beta[1,2,2]<-0 #no data
    beta[1,2,3]<-0 #no data
    beta[1,2,4]<-0 #no data
    beta[1,2,5]<-0 #no data (sample size 1)
    beta[2,1,1]~dnorm(BB[2,1],tauBB[2,1])
    beta[2,1,2]~dnorm(BB[2,1],tauBB[2,1])
    beta[2,1,3]~dnorm(BB[2,1],tauBB[2,1])
    beta[2,1,4]<-0 #no data
    beta[2,2,1]<-0 #no data
    beta[2,2,2]<-0 #no data
    beta[2,2,3]<-0 #no data
    beta[2,2,4]<-0 #no data
    beta[2,2,5]<-BB[2,2] #single subcategory
    beta[3,1,1]<-0 #no data
    beta[3,1,2]~dnorm(BB[3,1],tauBB[3,1])
    beta[3,1,3]~dnorm(BB[3,1],tauBB[3,1])
    beta[3,1,4]<-0 #no data
    beta[3,2,1]<-0 #no data
    beta[3,2,2]<-0 #no data
    beta[3,2,3]<-0 #no data
    beta[3,2,4]<-0 #no data
    beta[3,2,5]<-BB[3,2] #single subcategory
    
    BB[1,1]<-B[1]

    BB[1,2]<-0

    BB[2,1]~dnorm(B[2],tauB[1])
    tauBB[2,1]<-1/varBB[2,1]
    varBB[2,1]~dunif(0,1)

    BB[2,2]~dnorm(B[2],tauB[1])

    BB[3,1]~dnorm(B[3],tauB[2])
    tauBB[3,1]<-1/varBB[3,1]
    varBB[3,1]~dunif(0,1)

    BB[3,2]~dnorm(B[3],tauB[2])

B[1]~dnorm(0,0.1)

B[2]~dnorm(0,0.1)
tauB[1]<-1/varB[1]
varB[1]~dunif(0,1)

B[3]~dnorm(0,0.1)
tauB[2]<-1/varB[2]
varB[2]~dunif(0,1)

for(i in 1:W){ # SRE
iSRE[i]~dnorm(0,tau[R])
}

  for(i in 1:R){ 
tau[i]<-1/var[i]
var[i]~dunif(0,1)
  }

for(i in 1:K){    
tauAll[i]<-1/varAll[i]
varAll[i]~dunif(0,1)
}
    
} # end model

# initials
inits <- list(
  list(beta = c(rep(2, times=8)), #number of betas
       var = c(rep(0.5, times=R)), #one tau for all studies
       varBB=c(rep(0.5,times=2)), #number of varBB
       varB=c(rep(0.5,times=K-1)), #number of mechanisms, 3
       varAll=c(rep(1,times=K))),
  
  list(beta = c(rep(-2, times=8)), 
       var = c(rep(0.5, times=R)),
       varBB=c(rep(0.5,times=2)), 
       varB=c(rep(0.5,times=K-1)), 
       varAll=c(rep(1,times=K))),
  
  list(beta = c(rep(0, times=8)), 
       var = c(rep(0.5, times=R)),
       varBB=c(rep(0.5,times=2)), 
       varB=c(rep(0.5,times=K-1)), 
       varAll=c(rep(1,times=K)))
)

# parameters
parameters <- c("B", "BB", "beta", "natESavemean.h", "tau", "tauAll", "tauB", "tauBB", "varB", "varBB", "var")

# running model
es.simNatDist <- bugs(datamodel, inits, parameters, model.file = modelESNat, n.chains=3, n.burnin=25000, n.iter=75000, debug=F)
```

+ DIC value:
```{r echo=FALSE}
es.simNatDist[["DIC"]]
```

+ Trace plots

```{r echo=FALSE}
traplot(es.simNatDist, "beta")
traplot(es.simNatDist, "B")
traplot(es.simNatDist, "BB")
```

+ Model parameters 
```{r parameters 7, echo=FALSE}
x<-as.data.frame(es.simNatDist[["summary"]]) %>% rownames_to_column(., var = "rowname") %>% 
  select(rowname, mean, sd, `2.5%`, `97.5%`) %>% filter(grepl("^beta", rowname)|
                                                   grepl("^var", rowname)|
                                                     grepl("^B", rowname))
write.csv(x, file.path(parametersout, "ESNat-Disturbance.csv"), row.names = F)
x
```


+ Model fit:

```{r echo=FALSE}
x<-as.data.frame(es.simNatDist[["summary"]])
natESavemean.h <- x %>% select(mean) %>% rownames_to_column(., var = "rowname") %>% filter(grepl("^natESavemean.h", rowname)) %>% select(-rowname)
data3 <- cbind(natESavemean.h,natESavemean) 
colnames(data3) <- c("predicted", "estimated")
m1 <- lm(predicted~estimated, data = data3)
plot(predicted~estimated, data = data3, main = "ESNatmean\npredicted vs. estimated values")
abline(m1, col = 'red', lty = 'dashed', lwd = 1.5)
r2 = summary(m1)$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(r2)))
text(x = -1, y = 0.5, labels = mylabel)
```

```{r include=FALSE}
# data preparation for plot (native ES)

samplesize2 <- NatAllDataModel2 %>% 
  group_by(Mechanism) %>%
  add_tally() %>% ungroup() %>% mutate(Disturbance=NA, Type=NA) %>% 
  select(Mechanism, Disturbance, Type, n) %>% distinct() %>% arrange(Mechanism) %>% 
  mutate(index=c(1,5,11),
         size=14, # size font
         label=c("Propagule pressure", "Empty niches", "Low biotic resistance"),
         n=ifelse(index==1|index==11, n+1, n),shape=23,sizedot=6,treatment="a") # a=non significant 

samplesize3 <- NatAllDataModel2 %>% 
  group_by(Mechanism, Disturbance) %>%
  add_tally() %>% ungroup() %>% mutate(Type=NA) %>% 
  select(Mechanism, Disturbance, Type, n) %>% distinct() %>% arrange(Mechanism, Disturbance) %>%
  mutate(index=c(2,7,6,13,12),shape=23,sizedot=5,size=12,treatment=c("a","a","a","a","a"),
         label=c("Disturbance","Disturbance","No disturbance","Disturbance","No disturbance"))

samplesize <- NatAllDataModel2 %>% 
  group_by(Mechanism, Disturbance, Type) %>%
  add_tally() %>% ungroup() %>% select(Mechanism, Disturbance, Type, n) %>% distinct() %>% 
 arrange(Mechanism, Disturbance, Type) %>% 
  mutate(index=c(4,8,9,10,NA,15,16,NA),size=10,
         label=c("Human","Natural","Human","Both",NA,"Human","Both",NA),
         shape=23,sizedot=4,treatment=c("a","a","a","a",NA,"a","a",NA)) %>% 
  filter(!is.na(index)) %>% 
 rbind(samplesize2) %>% rbind(samplesize3) %>% mutate(n = paste0('(', n, ')')) 

ESNatEXCLUDED <- NatAllDataModelEXCLUDED %>% mutate(rowname=c("beta[1,1,1]","beta[3,1,1]"),name=c("beta","beta"),
                                                    `2.5%`=ESNatmean-(ESNatSD*2),
                                                    `97.5%`=ESNatmean+(ESNatSD*2),
                                                    n = paste0('(', 1, ')'),
                                                    index=c(3,14),size=10,shape=23,sizedot=5,treatment=c("b","b"),
                                                    label=c("Natural","Natural")) %>% 
  rename(mean=ESNatmean, sd=ESNatSD)
         
nat <- as.data.frame(es.simNatDist[["summary"]])
forestnat <- nat %>% 
  select(mean, sd, `2.5%`, `97.5%`) %>% 
  rownames_to_column(., var = "rowname") %>% 
  filter(!grepl("^natESavemean.h", rowname),
         !grepl("^tau", rowname),
         !grepl("^var", rowname),
         rowname != "deviance") %>% 
  separate(col=rowname, remove=F, into=c("name","Mechanism","Disturbance","Type")) %>% 
  mutate(Disturbance=as.numeric(Disturbance),
         Mechanism=as.numeric(Mechanism),
         Type=as.numeric(Type)) %>% 
  mutate_all(na_if,"") %>% 
  left_join(samplesize) %>% 
  filter(!is.na(index)) %>% 
  rbind(ESNatEXCLUDED) %>% 
  arrange(index) %>% 
  mutate(treatment2=ESNatcolor)
```

```{r include=FALSE}
# figure with invasive species and native community performance together (mechanism = propagule pressure)
PPnat <- forestnat %>% filter(Mechanism==1) %>% mutate(category="nat")

PPinv <- forestinv %>% filter(Mechanism==1) %>% mutate(category="inv",treatment=ifelse(treatment=="a","c","d"))

PP <- PPnat %>% rbind(PPinv,NA,NA,NA,NA) %>% mutate(rowname=ifelse(is.na(rowname),row_number(),rowname),
                                                    category=ifelse(rowname=="10","nat",
                                                            ifelse(rowname=="11","inv",
                                                            ifelse(rowname=="12","nat",
                                                            ifelse(rowname=="13","inv",category)))),
                                                    rowname=ifelse(rowname=="10","BB[1,2]",
                                                            ifelse(rowname=="11","beta[1,1,3]",
                                                            ifelse(rowname=="12","beta[1,1,3]",
                                                            ifelse(rowname=="13","beta[1,1,4]",rowname))))) %>%
  group_by(rowname) %>% mutate(index=cur_group_id(),index=ifelse(category=="nat", index+0.2, index),index=ifelse(rowname=="BB[1,2]", index-1,ifelse(rowname=="BB[1,1]", index+1,index))) %>% arrange(index)

 p1<-
  ggplot(data=PP, 
       aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`)) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  geom_errorbarh(data=PP, height=.1, size=1, color=PP$treatment2)+
  geom_point(data=PP, shape=PP$shape,size=PP$sizedot, color=PP$treatment2,aes(fill=treatment),stroke=1)+ 
  scale_y_continuous(name = "",
                     labels = c("Overall","No disturbance","Disturbance","Natural","Human","Both","Not reported","","","","","",""),
                     breaks=1:nrow(PP), trans="reverse")+
  scale_fill_manual(values = c('white', ESNatcolor, "white", ESInvcolor))+
  theme_bw()+
  ggtitle("Propagule pressure")+
  scale_x_continuous(expand = c(0.05,0.05))+
  coord_cartesian(ylim=c(7.1,0.95),xlim=c(-2.5,2.5),clip = "off")+
  xlab("Effect size (ES, mean + 95%PI)") +
  theme(text=element_text(size=13, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      axis.text.y = element_text(size=c(14,12,12,10,10,10,10,10,10),
                                 face=c("bold","plain","plain","plain","plain","plain","plain","plain","plain")),
      axis.text.x=element_text(size=10),
      panel.border = element_blank(),
      axis.title.y = element_blank(),
      plot.title = element_text(face = "bold",hjust = 0.5),
      plot.margin = margin(1,1,1,1),#defines margin so the graph does not cut off 
      axis.line = element_line(colour = "black"),
      legend.position = "none") +    #remove top and right panel boundaries
  geom_text_repel(aes(label = n), size=4, color=PP$treatment2)+
  annotate("text", x=-2.55, y=0.35, label="(a)", color="black", fontface = "bold", size=5)+
  ggh4x::force_panelsizes(rows = unit(4.5, "in"), cols = unit(4, "in"))

```

```{r include=FALSE}
# figure with invasive species and native community performance together (mechanism = biotic resistance)
BRnat <- forestnat %>% filter(Mechanism==3) %>% mutate(category="nat")

BRinv <- forestinv %>% filter(Mechanism==3) %>% mutate(category="inv",treatment=ifelse(treatment=="a","c","d"))

BR <- BRnat %>% rbind(BRinv) %>% arrange(rowname) %>% group_by(rowname) %>% mutate(index=cur_group_id(),index=ifelse(category=="nat", index+0.2, index),index=ifelse(rowname=="BB[3,2]", index-1,ifelse(rowname=="BB[3,1]", index+1,index))) %>% arrange(index)

p2<-
  ggplot(data=BR, 
       aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`)) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  geom_errorbarh(data=BR, height=.1, size=1, color=BR$treatment2)+
  geom_point(data=BR, shape=BR$shape,size=BR$sizedot, color=BR$treatment2,aes(fill=treatment),stroke=1)+ 
  scale_y_continuous(name = "",
                     labels = NULL,
                     breaks=1:nrow(BR), trans="reverse")+
  scale_fill_manual(values = c('white', ESNatcolor, 'white', ESInvcolor))+
  theme_bw()+
  scale_x_continuous(expand = c(0.05,0.05))+
  coord_cartesian(ylim=c(7.1,0.95),xlim=c(-1.6,1.6),clip = "off")+
  xlab("Effect size (ES, mean + 95%PI)") +
  ggtitle("Low biotic resistance")+
  theme(text=element_text(size=13, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      axis.text.y = element_text(size=c(14,12,12,10,10,10,10,10,10),
                                 face=c("bold","plain","plain","plain","plain","plain","plain","plain","plain")),
      axis.text.x=element_text(size=10),
      panel.border = element_blank(),
      axis.title.y = element_blank(),
      plot.title = element_text(face = "bold",hjust = 0.5),
      plot.margin = margin(1,1,1,10),#defines margin so the graph does not cut off (top,right,below,left)
      axis.line = element_line(colour = "black"),
      legend.position = "none") +    #remove top and right panel boundaries
  geom_text_repel(aes(label = n), size=4, color=BR$treatment2)+
  annotate("text", x=-1.65, y=0.32, label="(b)", color="black", fontface = "bold", size=5)+
  ggh4x::force_panelsizes(rows = unit(4.5, "in"), cols = unit(4, "in"))
```

```{r include=FALSE}
# figure with invasive species and native community performance together (mechanism = empty niches)
ENnat <- forestnat %>% filter(Mechanism==2) %>% mutate(category="nat")

ENinv <- forestinv %>% filter(Mechanism==2) %>% mutate(category="inv",treatment=ifelse(treatment=="a","c","d"))

EN <- ENnat %>% rbind(ENinv) %>% arrange(rowname) %>% group_by(rowname) %>% mutate(index=cur_group_id(),index=ifelse(category=="nat", index+0.2, index),index=ifelse(rowname=="BB[2,2]", index-1,ifelse(rowname=="BB[2,1]", index+1,index))) %>% arrange(index)

 p3<-
  ggplot(data=EN, 
       aes(y=index, x=mean, xmin=`2.5%`, xmax=`97.5%`)) +
  geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5, size=1)+
  geom_errorbarh(data=EN, height=.1, size=1, color=EN$treatment2)+
  geom_point(data=EN, shape=EN$shape,size=EN$sizedot, color=EN$treatment2,aes(fill=treatment),stroke=1)+ 
  scale_y_continuous(name = "",
                     labels=NULL,
                     breaks=1:nrow(EN), trans="reverse")+
  scale_fill_manual(values = c('white', "white", ESInvcolor))+
  theme_bw()+
  ggtitle("Empty niches")+
  scale_x_continuous(expand = c(0.05,0.05))+
  coord_cartesian(ylim=c(7.1,0.95),xlim=c(-1.3,1.8),clip = "off")+
  xlab("Effect size (ES, mean + 95%PI)") +
  theme(text=element_text(size=13, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      axis.text.y = element_text(size=c(14,12,12,10,10,10,10,10,10),
                                 face=c("bold","plain","plain","plain","plain","plain","plain","plain","plain")),
      axis.text.x=element_text(size=10),
      panel.border = element_blank(),
      axis.title.y = element_blank(),
      plot.title = element_text(face = "bold",hjust = 0.5),
      plot.margin = margin(1,1,1,10),#defines margin so the graph does not cut off 
      axis.line = element_line(colour = "black"),
      legend.position = "none") +    #remove top and right panel boundaries
  geom_text_repel(aes(label = n), size=4, color=EN$treatment2)+
  annotate("text", x=-1.35, y=0.32, label="(c)", color="black", fontface = "bold", size=5)+
  ggh4x::force_panelsizes(rows = unit(4.5, "in"), cols = unit(4, "in"))
```

```{r include=FALSE}
p1 <- ggplotGrob(p1)
p2 <- ggplotGrob(p2)
p3 <- ggplotGrob(p3)

# showing them all
png(file.path(graph.wd, "ESNatESInv_Dist.png"),width=13.8, height=5.5, res=300,units="in")
grid::grid.draw(cbind(p1,p2,p3)) 
dev.off()
```


# **SM: The World Map**

```{r coord objt, include=FALSE}
sites_coord <- data %>%
  rename(Latitude=Lat_m,
         Longitude=Long_m) %>% 
  select(StudyIDOriginal, Latitude, Longitude, Country) %>% #selects just a few of the columns
  distinct(StudyIDOriginal, .keep_all = TRUE) %>% #removes duplicates due to several data entries for the same
  filter(!is.na(Latitude)) # removes studies that the coordinates were not entered
sites_coord <- st_as_sf(sites_coord, coords = c("Longitude", "Latitude"), remove = FALSE, 
    crs = 4326, agr = "constant")
```

```{r world map, echo=FALSE}
world <- ne_countries(scale = "medium", returnclass = "sf") #creates the object with the map
ggplot(data = world) + #plot the object as a map
    geom_sf() + 
  coord_sf(expand = FALSE) + #this makes the coordinates to show up
    geom_point(data = sites_coord, aes(x = Longitude, y = Latitude), size = 4, #specifies the data
        shape = 21, fill = "darkred") + #specifies how the dots will show
  theme(panel.border = element_blank()) +
  theme(plot.margin = unit(c(0,0.5,0,0.5), "cm")) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
    xlab("Longitude") + ylab("Latitude") +
    ylim(-60,NA) + #defining the limits in the yaxis, NA to indicate that automatic upper limit
    ggtitle("World map of study sites included in the meta-analysis", 
            subtitle = paste0("(", length(unique(sites_coord$StudyIDOriginal)), " studies from ", length(unique(sites_coord$Country)), " countries)")) 
# saves
ggsave(file.path(graph.wd,paste(c("map", format(Sys.time(), "%m%d%Y"), ".png"), collapse="")), dpi=300, scale=2, width=7, height=3.2) #saves the last generated plot
```

# **SM: Funnel plots**
```{r include=FALSE}
# Invasive performance
glimpse(InvEstimate)

InvEstimateFP <- InvEstimate %>%
  mutate(ESInvSD2 = ESInvSD*2,
         Variance = 1/(ESInvSD^2))

plot1<-ggplot(InvEstimateFP) +
  geom_point(data=InvEstimateFP,
       aes(y=Variance, x=ESInvmean), size=2) +
  geom_errorbarh(data=InvEstimateFP, aes(y=Variance, xmax = ESInvmean + ESInvSD2, xmin = ESInvmean - ESInvSD2, height=5000)) +
  scale_y_continuous(limits = c(-5000, 170000), 
                     labels = function(x) format(x, scientific = TRUE)) + # yaxis as scientific notation
  geom_vline(xintercept=0, linetype="dashed", alpha=.5, size=1) +
  ylab("1/ESvariance") +
  xlab("ES (mean+2SD)") +
  ggtitle("(a) ES Invasive species\nperformance") +
    theme(text=element_text(size=10, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title.x = element_text(size = 10),
       plot.title = element_text(size=10),
      plot.margin = margin(2,20,2,2),#defines margin so the graph does not cut off 
      axis.line = element_line(colour = "black"),
      panel.border = element_blank())
```

```{r include=FALSE}
# Native Performance
glimpse(NatBioticEstimate)

NatBioticEstimateFP <- NatBioticEstimate  %>%
  mutate(ESNatSD2 = ESNatSD*2,
         Variance = 1/(ESNatSD^2))

plot2<-ggplot(NatBioticEstimateFP) +
  geom_point(data=NatBioticEstimateFP,
       aes(y=Variance, x=ESNatmean), size=2) +
  geom_errorbarh(data=NatBioticEstimateFP, aes(y=Variance, xmax = ESNatmean + ESNatSD2, xmin = ESNatmean - ESNatSD2, height=100000)) +
  scale_y_continuous(limits = c(-100000, 1400000), 
                     labels = function(x) format(x, scientific = TRUE)) + # yaxis as scientific notation
  geom_vline(xintercept=0, linetype="dashed", alpha=.5, size=1) +
  xlab("ES (mean+2SD)") +
  ggtitle("(b) ES Native community\nperformance") +
    theme(text=element_text(size=10, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title.x = element_text(size = 10),
      axis.title.y = element_blank(),
       plot.title = element_text(size=10),
      plot.margin = margin(2,20,2,2),#defines margin so the graph does not cut off 
      axis.line = element_line(colour = "black"),
      panel.border = element_blank())
```

```{r include=FALSE}
# Abiotic ES
glimpse(NatAbioticEstimate)

NatAbioticEstimateFP <- NatAbioticEstimate  %>%
  mutate(ESNatSD2 = ESNatSD*2,
         Variance = 1/(ESNatSD^2))

plot3<-ggplot(NatAbioticEstimateFP) +
  geom_point(data=NatAbioticEstimateFP,
       aes(y=Variance, x=ESNatmean), size=2) +
  geom_errorbarh(data=NatAbioticEstimateFP, aes(y=Variance, xmax = ESNatmean + ESNatSD2, xmin = ESNatmean - ESNatSD2, height=10000)) +
  scale_y_continuous(limits = c(-10000, 250000), 
                     labels = function(x) format(x, scientific = TRUE)) + # yaxis as scientific notation
  geom_vline(xintercept=0, linetype="dashed", alpha=.5, size=1) +
  xlab("ES (mean+2SD)") +
  ggtitle("(c) ES Abiotic") +
    theme(text=element_text(size=10, color="black"),
      panel.spacing = unit(1, "lines"),
      panel.grid = element_blank(),
      legend.position = "none",
      axis.title.x = element_text(size = 10),
      axis.title.y = element_blank(),
      plot.margin = margin(2,20,2,2),#defines margin so the graph does not cut off 
      axis.line = element_line(colour = "black"),
      plot.title = element_text(size=10),
      panel.border = element_blank())
```

```{r echo=FALSE}
par(mfrow = c(1, 3))
plot1
plot2
plot3
dev.off()
```


```{r warning=FALSE, include=FALSE}
p1 <- ggplotGrob(plot1)
p2 <- ggplotGrob(plot2)
p3 <- ggplotGrob(plot3)

# saves
grid::grid.newpage()
png(file.path(graph.wd, "FunnelPlots.png"),width=8, height=3, res=300,units="in")
grid::grid.draw(rbind(cbind(p1,p2, p3))) # allow graphs to align right
dev.off()
```
